---
title: "Portfoliomanagement and Financial Analysis - Assignment 4"
subtitle: "Submit until Monday 2019-10-07, 13:00"
author: "Amann, Matthias"
output: html_notebook
---

```{r load_packs}
pacman::p_load(tidyverse,tidyquant,PortfolioAnalytics,nloptr)
```

**Please** remember to put your assignment solutions in `rmd` format using **many** chunks and putting readable text in between, similar to my examples given in Research Methods and Assignment 1! Also, each student has to select his own set of 10 stocks having data available as of `2000-01-01`. Select by Sharpe-ratio, dominance or any other method (e.g. matching your first name).

For all exercises: Please use the Assignment-Forum to post your questions, I will try my best to help you along!

## Exercise 1: Rebalancing

Have a look at `vignette("ROI_vignette")` and the `optimize.portfolio.rebalancing` command. Use your dataset to compute 





```{r load_packs}
pacman::p_load(tidyverse,tidyquant,PortfolioAnalytics,nloptr)
```

**Please** remember to put your assignment solutions in `rmd` format using **many** chunks and putting readable text in between, similar to my examples given in Research Methods and Assignment 1! Also, each student has to select his own set of 10 stocks having data available as of `2000-01-01`. Select by Sharpe-ratio, dominance or any other method (e.g. matching your first name).

For all exercises: Please use the Assignment-Forum to post your questions, I will try my best to help you along!

## Exercise 1: Rebalancing

Have a look at `vignette("ROI_vignette")` and the `optimize.portfolio.rebalancing` command. Use your dataset to compute 

a) Mean-Return (Tangency portfolio)
b) Minimum-Variance
c) Maximum Quadratic Utility Portfolios

checking for a variety of constraints (constraints that can be computed with the `ROI`-solver) and different rebalancing periods (as well as rolling windows/training periods) to find, what might deliver you the best portfolios performance (use appropriate statistics to decide on that).

```{r}
library(timetk)

SP500 <- tq_index("SP500")
NASDAQ <- tq_exchange("NASDAQ")
NYSE <- tq_exchange("NYSE") 

stocks.selection <- SP500 %>% 
  inner_join(rbind(NYSE,NASDAQ) %>% select(symbol,last.sale.price,market.cap,ipo.year),by=c("symbol")) %>%
  filter(ipo.year<2000&!is.na(market.cap)) %>% 
  arrange(desc(weight)) %>% 
  slice(1:10)

stocks.selection
```

These are the returns of the selected stocks.

```{r}
stocks.returns <- stocks.selection$symbol %>%
    tq_get(get  = "stock.prices",
           from = "2000-01-01",
           to   = "2019-12-31") %>%
    group_by(symbol) %>%
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly")

stocks.returns
```

Stock returns as time series
````{r}
stocks.returns.xts <- stocks.returns%>%
                      subset( select = c(symbol,date, monthly.returns)) %>% 
                      pivot_wider(names_from = symbol, 
                                  values_from = monthly.returns) %>% 
                      tk_xts(date_var = date, silent = TRUE)

stocks.returns.xts
```

1A

Create portfolio object, add constraints, add objective
```{r}
portf_maxret <- portfolio.spec(assets=stocks.selection$symbol)%>%
  add.constraint(type="full_investment") %>%
  add.constraint(type="long_only") %>%
  add.objective(type="return", name="mean")

portf_maxret
```

Run the optimization
```{r}
opt_maxret <- optimize.portfolio(R=stocks.returns.xts, portfolio=portf_maxret,
                                 optimize_method="ROI", trace=TRUE)
opt_maxret
```

```{r}
 plot(opt_maxret, chart.assets=TRUE, main="Maximum Return", 
      xlim=c(0.05, 0.35), ylim=c(0,0.04))

```

```{r}
chart.RiskReward(opt_maxret,return.col="mean", risk.col="sd",
                 chart.assets=TRUE, 
                 xlim=c(0.05, 0.25),
                 main="Maximum Return")
```

```{r}
maxret.ef <- create.EfficientFrontier(R=stocks.returns.xts, 
                                       portfolio=portf_maxret, 
                                       type="mean-StdDev")

chart.EfficientFrontier(maxret.ef, match.col="StdDev", type="l")
```

Backtesting:  Quarterly rebalancing with 5 year training period
```{r}
bt_maxret <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_maxret,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period=60)
bt_maxret 
```

Backtesting:  Quarterly rebalancing with 5 year training period and 4 year rolling window
```{r}
bt_maxret2 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_maxret,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period= 60,
                                            rolling_window = 48)
bt_maxret2
```


1B

Create portfolio object, add constraint, add objective
```{r}
portf_minvar <- portfolio.spec(assets= stocks.selection$symbol) %>%
  add.constraint(type="full_investment") %>%
  add.constraint(type = "long_only") %>%
  add.objective(type="risk", name="var")
  
portf_minvar
```

Run the optimization
```{r}
opt_minvar <- optimize.portfolio(R = stocks.returns.xts,portfolio = portf_minvar,
                              optimize_method = "ROI", trace = TRUE)

opt_minvar
```

Plotting weights
```{r}
plot(opt_minvar, chart.assets=TRUE, main="Minimum Variance", 
      xlim=c(0.05, 0.35), ylim=c(0,0.05))
```

```{r}
chart.RiskReward(opt_minvar,return.col="mean", risk.col="sd",
                 chart.assets=TRUE, 
                 xlim=c(0.05, 0.25),
                 main="Minimum Variance")
```

Backtesting: Quarterly rebalancing with 5 year training period
```{r}
bt_minvar <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_minvar,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period=60)
bt_minvar
```

Backtesting: Quarterly rebalancing with 5 year training period and 4 year rolling window
```{r}
bt_minvar2 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_minvar,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period=60,
                                            rolling_window = 48)
bt_minvar2
```
1C

```{r}
portf_maxqua <- portfolio.spec(assets= stocks.selection$symbol) %>%
  add.constraint(type = "full_investment") %>%
  add.constraint(type = "long_only")
portf_maxqua
```

Create return objective, risk aversion objective and combine
```{r}
ret_obj <- return_objective(name="mean")

var_obj <- portfolio_risk_objective(name="var", risk_aversion=1)

qu_obj <- list(ret_obj, var_obj)

qu_obj
```

Run the optimization
```{r}
opt_qu <- optimize.portfolio(R=stocks.returns.xts, portfolio=portf_maxqua,
                             objectives=qu_obj,
                             optimize_method="ROI",
                             trace=TRUE)
opt_qu
```

```{r}
plot(opt_qu, chart.assets=TRUE, main="Maximum Quadratic Utility", 
      xlim=c(0, 0.35), ylim=c(0,0.05))
```

```{r}
chart.RiskReward(opt_qu,return.col="mean", risk.col="sd",
                 chart.assets=TRUE, 
                 xlim=c(0.05, 0.25),
                 main="Maximum Quadratic Utility")
```

Backtesting: Quarterly rebalancing with 5 year training period
```{r}
bt_qu <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60)
bt_qu
```


Backtesting: Quarterly rebalancing with 5 year training period
```{r}
bt_qu2 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60,
                                        rolling_window = 48)
bt_qu2
```

### With risk aversion = 5
Create risk aversion objective and combine with return objective
```{r}
var_obj2 <- portfolio_risk_objective(name="var", risk_aversion=5)

qu_obj2 <- list(ret_obj, var_obj2)

qu_obj2
```

Run the optimization
```{r}
opt_qu2 <- optimize.portfolio(R=stocks.returns.xts, portfolio=portf_maxqua,
                             objectives=qu_obj2,
                             optimize_method="ROI",
                             trace=TRUE)
opt_qu2
```

```{r}
plot(opt_qu2, chart.assets=TRUE, main="Maximum Quadratic Utility", 
      xlim=c(0, 0.35), ylim=c(0,0.05))
```


```{r}
chart.RiskReward(opt_qu2,return.col="mean", risk.col="sd",
                 chart.assets=TRUE, 
                 xlim=c(0.05, 0.25),
                 main="Maximum Quadratic Utility")
```

Backtesting: Quarterly rebalancing with 5 year training period
```{r}
bt_qu3 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj2,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60)
bt_qu
```


Backtesting: Quarterly rebalancing with 5 year training period
```{r}
bt_qu4 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj2,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60,
                                        rolling_window = 48)
bt_qu2
```


### Comparing Portfolio Performances with CAPM

weights of the portfolios
```{r}
wts_maxret <- c(0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
wts_minvar <- c(0.0344, 0.0821, 0, 0, 0.006, 0, 0.2965, 0.0021, 0.5143, 0.0645)
wts_qu <- c(0.5127, 0, 0.1358, 0.1965, 0.0693, 0, 0.0857, 0, 0, 0)
wts_qu2 <- c(0.1738, 0.0173, 0.0047, 0.0217, 0.0706, 0, 0.3064, 0, 0.4055, 0)

wts_maxret
wts_minvar
wts_qu 
wts_qu2
```

Portfolio Returns
```{r}
stocks.returns

portfolio.returns.maxret <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_maxret, 
                 col_rename  = "Ra")

portfolio.returns.maxret

portfolio.returns.minvar <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_minvar, 
                 col_rename  = "Ra")

portfolio.returns.minvar

portfolio.returns.qu <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_qu, 
                 col_rename  = "Ra")

portfolio.returns.qu

portfolio.returns.qu2 <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_qu2, 
                 col_rename  = "Ra")

portfolio.returns.qu2
```

Baseline Returns S&P 500
```{r}
baseline.returns <- "^GSPC" %>%
    tq_get(get  = "stock.prices",
           from = "2000-01-01",
           to   = "2019-12-31") %>%
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly", 
                 col_rename = "Rb")

baseline.returns
```


Merging the datasets together to compute CAPM
```{r}
RaRb.maxret <- left_join(portfolio.returns.maxret, 
                                   baseline.returns,
                                   by = "date")

RaRb.minvar <- left_join(portfolio.returns.minvar, 
                                   baseline.returns,
                                   by = "date")

RaRb.qu <- left_join(portfolio.returns.qu, 
                                   baseline.returns,
                                   by = "date")

RaRb.qu2 <- left_join(portfolio.returns.qu2, 
                                   baseline.returns,
                                   by = "date")

RaRb.maxret
RaRb.minvar
RaRb.qu 
RaRb.qu2 
```

CAPM MaxRet Portfolio
```{r}
RaRb.maxret %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```


CAPM MinVar Portfolio
```{r}
RaRb.minvar %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```

CAPM Quadratic Utility Portfolio risk_aversion=1
```{r}
RaRb.qu %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```

CAPM Quadratic Utility Portfolio risk_aversion=5
```{r}
RaRb.qu2 %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```

```{r}
portfolio.returns.maxret %>%
    ggplot(aes(x = date, y = Ra)) +
    geom_bar(stat = "identity", fill = palette_light()[[1]]) +
    labs(title = "MaxRet Portfolio Returns",
         x = "", y = "Monthly Returns") +
    geom_smooth(method = "lm") +
    theme_tq() +
    scale_color_tq() +
    scale_y_continuous(labels = scales::percent)
```

```{r}
portfolio.returns.minvar %>%
    ggplot(aes(x = date, y = Ra)) +
    geom_bar(stat = "identity", fill = palette_light()[[1]]) +
    labs(title = "MinVar Portfolio Returns",
         x = "", y = "Monthly Returns") +
    geom_smooth(method = "lm") +
    theme_tq() +
    scale_color_tq() +
    scale_y_continuous(labels = scales::percent)
```

```{r}
portfolio.returns.qu %>%
    ggplot(aes(x = date, y = Ra)) +
    geom_bar(stat = "identity", fill = palette_light()[[1]]) +
    labs(title = "Quadratic Utility Portfolio Returns (Riskaversion=1)",
         x = "", y = "Monthly Returns") +
    geom_smooth(method = "lm") +
    theme_tq() +
    scale_color_tq() +
    scale_y_continuous(labels = scales::percent)
```


```{r}
portfolio.returns.qu2 %>%
    ggplot(aes(x = date, y = Ra)) +
    geom_bar(stat = "identity", fill = palette_light()[[1]]) +
    labs(title = "Quadratic Utility Portfolio Returns (Riskaversion=5)",
         x = "", y = "Monthly Returns") +
    geom_smooth(method = "lm") +
    theme_tq() +
    scale_color_tq() +
    scale_y_continuous(labels = scales::percent)
```

















## Exercise 2: Custom moments function

Check `vignette("custom_moments_objectives")` to implement a variety of robust covariance matrix estimates (see `?MASS::cov.rob`, `?PerformanceAnalytics::ShrinkageMoments` and maybe `?PerformanceAnalytics::EWMAMoments` - the latter one only for backtesting) for the minimum variance and quadratic utility portfolios. Plot the different Efficient frontiers, optimal portfolios and weights and visualize the different covariances. Also make yourselves comfortable with cleaning outliers from your timeseries via `return.Clean()`.

```{r}
library(parallel)
library(PortfolioAnalytics)
library(DEoptim)
library(MASS)
```


Construct initial portfolio with basic constraints.
```{r}
init.portf <- portfolio.spec(assets=stock_selection$symbol)
init.portf <- add.constraint(portfolio=init.portf, type="full_investment")
init.portf <- add.constraint(portfolio=init.portf, type="long_only")
init.portf
```

Add standard deviation as an objective (minimal variance portfolio)
  
```{r}
SD.portf.min <- add.objective(portfolio=init.portf, type="risk", name="StdDev")
```

Portfolio with standard deviation as an objective

```{r}
SD.portf.maxq <- add.objective(portfolio=init.portf, type="return", name="mean")
SD.port.maxq <- add.objective(portfolio=SD.portf, type="risk", name="var", risk_aversion=0.25)
```




function to estimate covariance matrix with cov.rob for minimum variance

```{r}
sigma.robust <- function(R){
      require(MASS)
      out <- list()
      set.seed(1234)
      out$sigma <- cov.rob(R, method="mcd")$cov
      return(out)
      }
sigmarob<-sigma.robust(stock_returns.xts)
sigmarob
```

function to estimate covariance matrix with ShrinkageMoments for minimum variance
```{r}
sigma.robust.shrink <- function(R){
    targets <- c(1,3,4)
    f <- rowSums(stock_returns.xts)
    out <- list()
    out$sigmashrink <- M2.shrink(stock_returns.xts, targets, f)$M2sh
    return(out)}
      
sigma.shrink <- sigma.robust.shrink(stock_returns.xts)
sigma.shrink
```

**Optimize portfolios**
```{r message=FALSE, warning=FALSE}

opt.sd.minv <- optimize.portfolio(stock_returns.xts, SD.port.minv, optimize_method="ROI", momentFUN="sigma.robust", trace = TRUE)

opt.sd.maxq <- optimize.portfolio(stock_returns.xts, SD.port.maxq, optimize_method="ROI", momentFUN="sigma.robust", trace = TRUE)

opt.sd.minv.shrink <- optimize.portfolio(stock_returns.xts, SD.port.minv, optimize_method="ROI", momentFUN="sigma.robust.shrink", trace = TRUE)

opt.sd.maxq.shrink <- optimize.portfolio(R=stock_returns.xts, portfolio=SD.port.maxq, optimize_method="ROI", momentFUN="sigma.robust.shrink", trace = TRUE)


```

Here we extract the weights and compute the portfolio standard deviation to verify that the
the robust estimate of the covariance matrix was used in the optimization
```{r}
weights <- extractWeights(opt.sd.minv)
  sigmarob <- sigma.robust(R=stock_returns.xts)
  sqrt(t(weights) %*% sigma %*% weights)
  extractObjectiveMeasures(opt.sd)$StdDev
```


**Plot the covariance matrix from cov.rob**
```{r echo = FALSE}
chart.Correlation(sigma, histogram = TRUE)
```

**Plot the covariance matrix from shrink**
```{r echo = FALSE}
chart.Correlation(sigma.shrink, histogram = TRUE)
```


**Plot the Portfolios**
```{r echo = FALSE}
plot(opt.sd.minv, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02))


plot(opt.sd.minv.shrink, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02))

plot(opt.sd.maxq, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.05))


plot(opt.sd.maxq.shrink, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.05))
```

**Chart Efficient Frontiert for the minimum variance Portfolio**
```{r echo = FALSE}
prt_eff_minv <- create.EfficientFrontier(R=stock_returns.xts, portfolio=SD.port.minv, type="mean-StdDev", match.col = "StdDev")
chart.EfficientFrontier(prt_eff_minv, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)
chart.EF.Weights(prt_eff_minv, colorset=rainbow(n = length(stock_selection)), match.col="StdDev", cex.lab = 1, main = "StdDev")
```


**Chart Efficient Frontiert for the quadratic utility Portfolio**
```{r echo = FALSE}
prt_eff_maxq <- create.EfficientFrontier(R=stock_returns.xts, portfolio=SD.port.maxq, type="mean-StdDev", match.col = "StdDev")
chart.EfficientFrontier(prt_eff_maxq, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)
chart.EF.Weights(prt_eff_maxq, colorset=rainbow(n = length(stock_selection)), match.col="StdDev", cex.lab = 1, main = "StdDev")
```




























## Exercise 3: Regime Switching

Have a look at `demo(regime_switching)` and estimate and rebalance portfolios based on 2/3 regimes. Can you plot the regimes over time?

```{r}
demo(regime_switching)
```



```{r load_packs}
pacman::p_load(tidyverse,tidyquant,PortfolioAnalytics,nloptr)
```




```{r}
# Load package and data.
library(PortfolioAnalytics)
```


```{r}
# Get monthly stock returns from the S&P500
monthly_returnsSP500 <- "^GSPC" %>%
  tq_get(get = "stock.prices", from = "2000-01-01", to = "2019-08-31") %>%
  tq_transmute(adjusted, periodReturn, period = "monthly", col_rename = "returns SP500")
monthly_returnsSP500
```

```{r}
# Calculate the rolling mean monthly

rollmeanSP500 <- rollmean(monthly_returnsSP500[, "returns SP500"], 2)
rollmeanSP500

```

```{r}
vector <- c(rollmeanSP500)

#2=good economy, 1=bad economy
regime1or2 <-as.numeric(vector>0)+1
regime1or2
```


```{r}
SP500dates <- select(monthly_returnsSP500,date)


#regime 1 is bad market phase (1) and regime 2 is good market phase (2)
data_frame <- data.frame(SP500dates[2:236,], regime1or2)
data_frame





```

```{r}
regime_xts <-data_frame %>%
  timetk::tk_xts(silent = TRUE)
regime_xts
```


```{r}
stockselection <- c("ABCB", "AAPL", "ACLS", "ADBE", "ADTN", "AEHR", "AEIS", "AHPI", "AKAM", "AMZN")

# Get the prices of the stocks to transmute it to returns

stock.prices <- stockselection %>%
  tq_get(get  = "stock.prices", from = "2000-01-01",to = "2018-08-31") %>%
  group_by(symbol)

# Create monthly returns
stock.returns.monthly <- stock.prices %>%  
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns"
               )

## Make a tibble with dates and returns for all stocks
## Make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts (necessary for Portfolioanalytics)

R <- pivot_wider(stock.returns.monthly,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>% 
  timetk::tk_xts(date_var = date, silent = TRUE)



colnames(R) <- c("ABCB", "AAPL", "ACLS", "ADBE", "ADTN", "AEHR", "AEIS", "AHPI", "AKAM", "AMZN")
funds <- colnames(R)



R %>% head()
```


```{r}
# Construct portfolio for regime 1 - bad economy.
## Here, the first regime is considered with a risk approach (Mean-ES portfolio and other constraints)   --> we optimize ES

### Es = Conditional Value at risk: considers losses that exceed the value-at-risk and determines their average amount. 

port1 <- portfolio.spec(funds)
port1 <- add.constraint(port1, "weight_sum", min_sum=0.99, max_sum=1.01)
port1 <- add.constraint(port1, "box", min=0.05, max=0.5)
port1 <- add.objective(port1, type="risk", name="ES", arguments=list(p=0.9))
port1 <- add.objective(port1, type="risk_budget", name="ES", 
                       arguments=list(p=0.9), max_prisk=0.5)

```


```{r}
## Construct portfolio for regime 2 - good economy.
### Here regime 2 is a regime based on standard investment with volatility - here we used the standard deviation --> we optimize Stdev

port2 <- portfolio.spec(funds)
port2 <- add.constraint(port2, "weight_sum", min_sum=0.99, max_sum=1.01)
port2 <- add.constraint(port2, "box", min=0, max=0.6)
port2 <- add.objective(port2, type="risk", name="StdDev")
port2 <- add.objective(port2, type="risk_budget", name="StdDev", max_prisk=0.5)
```


```{r}
# Combine the portfolios.
portfolios <- combine.portfolios(list(port1, port2))


# Now we construct the regime model and corresponding portfolios to use for each regime.

## we merge the portfolios and the regimes (because we cannot merge every single portfolio with the regimes)

regime.port <- regime.portfolios(regime_xts, portfolios)
regime.port
```
```{r}
### This optimization should result in out portfolio for regime 2
opt1 <- optimize.portfolio(R[1:(nrow(R)-1)], regime.port, 
                           optimize_method="DEoptim", 
                           search_size=2000, 
                           trace=TRUE)
```
```{r}
opt1
opt1$regime
```


```{r}
### This optimization should result in out portfolio for regime 1
opt2 <- optimize.portfolio(R[1:(nrow(R)-1)], regime.port, 
                           optimize_method="DEoptim", 
                           search_size=2000, 
                           trace=TRUE)
```
```{r}
opt2
opt2$regime
```


```{r}
# We can extract which regime portfolio we optimized with at each rebalance date.  

## If there are structural changes in the data series, maybe a date fits better in the other regime now
lapply(opt.rebal$opt_rebalancing, function(x) x$regime)
```


```{r}
## Extract the optimal weights at each rebalance date.
wt <- extractWeights(opt.rebal)
wt
```


```{r}
## Extract the objective measures.
obj <- extractObjectiveMeasures(opt.rebal)
str(obj)
obj
```



```{r}
## Extract the stats.
xt <- extractStats(opt.rebal)
str(xt)
```


```{r}
### Note that this returns a list of N elements for N regimes. We may have different objectives and/or a different number of objectives which makes returning a single xts object difficult/

# Extract the optimal weights at each rebalance date.
chart.Weights(opt.rebal, colorset=rainbow10equal)

wt
```

```{r}
# Chart the risk contribution for regime 1

chart.RiskBudget(opt.rebal, match.col="ES", risk.type="percentage", 
                 regime=1, colorset=rainbow10equal)

opt2
```


```{r}
# Chart the risk contribution for regime 2
chart.RiskBudget(opt.rebal, match.col="StdDev", risk.type="percentage", 
                 regime=2, colorset=rainbow10equal)

opt1
```

```{r}
# Chart the risk contribution for regime 2
chart.RiskBudget(opt.rebal, match.col="StdDev", risk.type="percentage", 
                 regime=2, colorset=rainbow10equal)

opt1
```




















## Exercise 4: Single Index-Model

Now we are going to estimate the Portfolio Input Parameters with the Single-Index Model. Use your ten assets and additionally choose the S&P500 as index (same returns etc).

a) Regress all stocks on the index. Show alpha, beta and residual variance. Calculate systematic and firm-specific risk. Are there any significant alphas? (You should double check with the appropriate `PerformanceAnalytics` Functions)


b) Extract the betas and calculate systematic and unsystematic risk, derive the whole covariance matrix. To do this you can use _CH15_Factor_Modfels_for_Asset_Returns.pdf (15.3.1)_ and the code
implemented in the function sharpeFactorEstimator that you find [here](http://financewithr.blogspot.com/2013/06/portfolio-optimization-using-single.html) (please do not just copy everything, but try to understand what you are doing, e.g. check why and if G.hat has the same values as found by the multivariate regression).


c) Now use the _custom-moments_ functions from Exercise 2 to implement the single-factor model into the portfolio optimization framework and plot the efficient frontier using the parameters estimated by the single factor model next to the EF of the full-covariance model. Calculate MVP, TP etc. and work out the differences in weights, portfolio return and portfolio risk.

#Load packages

```{r}
install.packages("fEcofin", repos="http://R-Forge.R-project.org")

library(fEcofin)
library(fPortfolio)

```


Now we are going to estimate the Portfolio Input Parameters with the Single-Index Model. Use your ten assets and additionally choose the S&P500 as index (same returns etc).

```{r}

stockselection_4 <- c("AAPL", "MSFT", "AMZN", "PFE", "JPM", "CSCO", "XOM", "BAC", "NKE", "WMT", "^GSPC")

#Get the prices of the stocks
stock.prices_4 <- stockselection_4 %>%
  tq_get(get  = "stock.prices", from = "2000-01-01",to = "2019-09-17") %>%
  group_by(symbol)
stock.returns.monthly_4 <- stock.prices_4 %>%  
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns")

stock.returns.monthly_xts_4 <- pivot_wider(stock.returns.monthly_4,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>%
  tk_xts(date_var = date, silent = TRUE)

```

a) Regress all stocks on the index. Show alpha, beta and residual variance. Calculate systematic and firm-specific risk. Are there any significant alphas? (You should double check with the appropriate `PerformanceAnalytics` Functions)

```{r}

#Regress all stocks on the index
alpha.Stocks <- CAPM.alpha(Ra = stock.returns.monthly_xts_4[,-11], Rb = stock.returns.monthly_xts_4[,11], Rf = 0)
beta.Stocks <- CAPM.beta(Ra = stock.returns.monthly_xts_4[,-11], Rb = stock.returns.monthly_xts_4[,11], Rf = 0)
StdDev.Index <- StdDev(R = stock.returns.monthly_xts_4[,11],
                   clean = "none",
                   method = "pearson")

lm(stock.returns.monthly_xts_4[,-11] ~ stock.returns.monthly_xts_4[,11])


plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 1], main = "Apple Regression", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 1] ~ stock.returns.monthly_xts_4[, 11])))
plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 2], main = "Microsoft Regression", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 2] ~ stock.returns.monthly_xts_4[, 11])))
plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 3], main = "Amazon Regression", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 3] ~ stock.returns.monthly_xts_4[, 11])))
plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 4], main = "Pfizer Regression", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 4] ~ stock.returns.monthly_xts_4[, 11])))
plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 5], main = "JPMorgan Regression", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 5] ~ stock.returns.monthly_xts_4[, 11])))
plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 6], main = "Cisco Regression", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 6] ~ stock.returns.monthly_xts_4[, 11])))
plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 7], main = "ExxonMobile", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 7] ~ stock.returns.monthly_xts_4[, 11])))
plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 8], main = "Bank of America", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 8] ~ stock.returns.monthly_xts_4[, 11])))
plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 9], main = "Nike Regression", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 9] ~ stock.returns.monthly_xts_4[, 11])))
plot.default(x = stock.returns.monthly_xts_4[, 11], y = stock.returns.monthly_xts_4[, 10], main = "Walmart Regression", xlab = "Index Returns", ylab = "Stock Returns", abline(lm(stock.returns.monthly_xts_4[, 10] ~ stock.returns.monthly_xts_4[, 11])))

#Calculate systematic (Market-Specific) Risk by mulitplying Variance (StdDev^2) of the S&P500 and the Beta^2 of each stock
sys.risk <- SystematicRisk(Ra = stock.returns.monthly_xts_4[,-11], Rb = stock.returns.monthly_xts_4[,11], Rf = 0)
sys.risk

#Calculate Firm-specific Risk / Residual Variance
firm.specific.risk <- SpecificRisk(Ra = stock.returns.monthly_xts_4[,-11], Rb = stock.returns.monthly_xts_4[,11], Rf = 0)
firm.specific.risk

#Summary
summary.SFM <- table.SFM(Ra = stock.returns.monthly_xts_4[,-11], Rb = stock.returns.monthly_xts_4[,11], scale = NA, Rf = 0, digits = 6)
summary.SFM

```

b) Extract the betas and calculate systematic and unsystematic risk, derive the whole covariance matrix. To do this you can use _CH15_Factor_Models_for_Asset_Returns.pdf (15.3.1)_ and the code
implemented in the function sharpeFactorEstimator that you find [here](http://financewithr.blogspot.com/2013/06/portfolio-optimization-using-single.html) (please do not just copy everything, but try to understand what you are doing, e.g. check why and if G.hat has the same values as found by the multivariate regression).

```{r}


#Calculate Beta of Portfolio by average each stocks beta
beta.portfolio <- mean(beta.Stocks)
beta.portfolio

#Calculate systematic (Market-Specific) Risk of portfolio
sys.risk.portfolio <- mean(sys.risk)
sys.risk.portfolio

#Calculate unsystematic risk by calculating the mean of the firm-specific risk
unsys.risk.portfolio <- mean(firm.specific.risk)
unsys.risk.portfolio

#Calculate Covariance Matrix
stock.returns.monthly_data <- as.data.frame((stock.returns.monthly_xts_4))
data(stock.returns.monthly_data)
returns <- as.timeSeries(stock.returns.monthly_data[,-11])
factors <- as.vector(as.timeSeries(stock.returns.monthly_data)[,11])
names(data)
data <- returns
attr(data, "factors") <- factors
nScenarios <- nrow(data)
X.mat <- cbind(rep(1, times=nScenarios), factors)
G.hat <- solve(qr(X.mat), data) #G.hat is alpha
beta.hat <- G.hat[2, ] #is beta
eps.hat <- data - X.mat %*% G.hat
diagD.hat <- diag(crossprod(eps.hat) / (nScenarios-2))
cov.si = var(factors)*(beta.hat%o%beta.hat) + diag(diagD.hat)
cov.si

```

c) Now use the _custom-moments_ functions from Exercise 2 to implement the single-factor model into the portfolo optimization framework and plot the efficient frontier using the parameters estimated by the single factor model next to the EF of the full-covariance model.

```{r}

#Function to implement single-factor model into the portfolio optimization framework
data(stock.returns.monthly_xts_4)
returns <- as.timeSeries(stock.returns.monthly_xts_4)
names(data)
data <- returns[, -c(11)]
factors <- returns[, 11]
attr(data, "factors") <- factors
# Sharpe's Single Index Factor Model:
sharpeFactorEstimator <- 
function(x, spec=NULL, ...)
{
    # Sharpe Single Index Model:
    data <- getDataPart(x)
    factors <- attr(x, "factors")
    nScenarios <- nrow(data)
    X.mat <- cbind(rep(1, times=nScenarios), factors)
    G.hat <- solve(qr(X.mat), data)
    beta.hat <- G.hat[2, ]
    eps.hat <- data - X.mat %*% G.hat
    diagD.hat <- diag(crossprod(eps.hat) / (nScenarios-2))
    mu <- G.hat[1, ] + G.hat[2, ] * colMeans(factors)  
    Sigma <- var(factors)[[1]] * (beta.hat %o% beta.hat) + diag(diagD.hat)
    
    # Return Value:
    list(mu = mu, Sigma = Sigma)
}
spec <- portfolioSpec()
setEstimator(spec) <- "sharpeFactorEstimator"
sharpe <- portfolioFrontier(data, spec)

#Chart the efficient frontier using the parameters estimated by the single factor model
sharpe_1 <- portfolioFrontier(data)
tailoredFrontierPlot(sharpe_1)
points(frontierPoints(sharpe), col = "steelblue")


#Chart Efficient Frontier minimum variance
chart.EfficientFrontier(prt_eff_minv, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)

```

Calculate MVP, TP etc. and work out the differences in weights, portfolio return and portfolio risk.

```{r}

#weights, portfolio return and portfolio risk of MVP
opt.sd.minv.shrink
plot(opt.sd.minv.shrink, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02)) 


#weights, portfolio return and portfolio risk of QUP
opt.sd.maxq.shrink
plot(opt.sd.maxq.shrink, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(-0.08,0.05))


#weights, portfolio return and portfolio risk of tangency portfolio (Highest risk/return ratio)
weight_tp_sharpe <- sharpe_1@portfolio@portfolio[["weights"]][26, ]
return_tp_sharpe <- sharpe_1@portfolio@portfolio[["targetReturn"]][26, ]
risk_tp_sharpe <- sharpe_1@portfolio@portfolio[["targetRisk"]][26, ]

weight_tp_sharpe
return_tp_sharpe
risk_tp_sharpe

```