---
title: "Portfoliomanagement and Financial Analysis - Assignment 4"
subtitle: "Submit until Monday 2019-10-07, 13:00"
author: "Matt, Sebastian"
output: html_notebook
---

```{r load_packs}
pacman::p_load(tidyverse,tidyquant,PortfolioAnalytics,nloptr)
```

**Please** remember to put your assignment solutions in `rmd` format using **many** chunks and putting readable text in between, similar to my examples given in Research Methods and Assignment 1! Also, each student has to select his own set of 10 stocks having data available as of `2000-01-01`. Select by Sharpe-ratio, dominance or any other method (e.g. matching your first name).

For all exercises: Please use the Assignment-Forum to post your questions, I will try my best to help you along!


## Exercise 1: Rebalancing

*Have a look at `vignette("ROI_vignette")` and the `optimize.portfolio.rebalancing` command. Use your dataset to compute* 

*a) Mean-Return (tangency portfolio)*
*b) Minimum-Variance*
*c) Maximum Quadratic Utility Portfolios*

*Checking for a variety of constraints (constraints that can be computed with the `ROI`-solver) and different rebalancing periods (as well as rolling windows/training periods) to find, what might deliver you the best portfolios performance (use appropriate statistics to decide on that).*

### Get the packages 

```{r}
# Load the necessary packages vignette("ROI_vignette")
suppressMessages(library(PortfolioAnalytics))
suppressMessages(library(foreach))
suppressMessages(library(iterators))
suppressMessages(library(ROI))
suppressMessages(library(ROI.plugin.quadprog))
suppressMessages(library(ROI.plugin.glpk))
```

### Personal Stock Selection

First I need my stocks!
I choose the following personal stockselection:

* **MO    =     Altria Group Rg (Philip Morris Companies)       (Sektor: Basiskonsumgüt,    Industrie: Tabak                                   )**
* **MDLZ  =     Mondelez International Inc.                     (Sektor: Basiskonsumgüt,    Industrie: Nahrungsmittel                          )**
* **PEP   =     PepsiCo Inc. (PEP)                              (Sektor: Basiskonsumgüt,    Industrie: Getränke                                )**
* **PM    =     Philip Morris International Inc.                (Sektor: Basiskonsumgüt,    Industrie: Tabak                                   )**
* **KHC   =     The Kraft Heinz Company                         (Sektor: Basiskonsumgüt,    Industrie: Nahrungsmittel                          )**
* **T     =     AT&T Inc.                                       (Sektor: Comm. Ser.,        Industrie: Integrated Telecommunication Services   )**
* **WMT   =     Walmart                                         (Sektor: Consumer Stap.,    Industrie: Hypermarkets & Super Centers            )**
* **MSFT  =     Microsoft Corp.                                 (Sektor: Information Tech,  Industrie: Systems Software                        )**
* **BAC   =     Bank of America Cor                             (Sektor: Financials,        Industrie: HDiversified Bankss                     )**
* **PG    =     Procter & Gamble                                (Sektor: Consumer Stap.,    Industrie: Personal Products                       )**


("MO", "MDLZ", "PEP", "PM", "KHC", "T","WMT", "MSFT","BAC","PG")


Attention not all stocks have datas up to the chosen date. So ATTENTION when choosing them. Also pay atention to the leng of some vectors, when you extend or reduce the size of the stock selection. Next key point is the input in the function "optimize.portfolio.rebalancing()"

* optimize_method = "ROI"
* rebalance_on = "month"
* training_period = 10

Look for the right "time frame" and the numbers of training period (part of the task!!!)


```{r}
stockselection_1 <- c("MO", "MDLZ", "PEP", "PM", "KHC", "T","WMT", "MSFT","BAC","PG")
# Get the prices of the stocks 
stock.prices <- stockselection_1 %>%
  tq_get(get  = "stock.prices", 
         from = "2015-07-06",    #First trade day of KHC
         to = Sys.Date( )) %>% 
  
  group_by(symbol)
stock.prices %>% slice(1:2) 
# Output the first two entries of each stock!
stock.prices %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  ggtitle("Price chart for all stocks - all in one")
# Plotting the stock prices in each frame
stock.prices %>%
  ggplot(aes(x = date, y = adjusted)) +
  geom_line() +
  facet_wrap(~symbol, scales = "free_y") +
  theme_classic() +
  labs(x = "Date", y = "Price") +
  ggtitle("Price chart all stocks - in each frame") 
```

```{r}
# Create monthly returns by the tq_transmute() = adds new variables to an existing tibble;
stock.returns.monthly <- stock.prices %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns"
               )
# Output the first two entries of each stock!
stock.returns.monthly %>% slice(1:2) 
# Make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts
stock.returns.monthly_xts <- pivot_wider(stock.returns.monthly,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>%
   tk_xts(date_var = date, silent = TRUE)
# Output the first entries (simple returns from adjusted prices) of each stock!
stock.returns.monthly_xts[1]
# Plotting a performance summary chart 
charts.PerformanceSummary(stock.returns.monthly_xts, 
                          main="Performance summary")
```

### **a) Mean-Return**
***Maximizing Mean Return Portfolio***
### Theorie: Maximizing Mean Return with ROI

The objective to maximize mean return is a linear problem of the form: $maximize_w=µˆ'\cdot w$
Where µˆ is the estimated mean asset returns and w is the set of weights. Because this is a linear problem, it is well suited to be solved using a linear programming solver. For these types of problems, PortfolioAnalytics uses the ROI package with the glpk plugin.

*Creating a portfolio object with the function "portfolio.spec()"*
```{r}
# Maximazing Mean Return 
# Create a portfolio object
portf_maxret <- portfolio.spec(assets=stockselection_1)
# Creat min and max weights for the portfolio 
min_w <- c(0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02)
max_w <- c(0.60, 0.60, 0.60, 0.60, 0.60, 0.60, 0.60, 0.60, 0.60, 0.60)
# Add constraints to the portfolio object 
portf_maxret <- add.constraint(portfolio=portf_maxret, type="full_investment")
portf_maxret <- add.constraint(portfolio=portf_maxret, type="box", 
                               min=min_w, 
                               max=max_w)
# Add objective to the portfolio object
portf_maxret <- add.objective(portfolio=portf_maxret, type="return", name="mean") 
```


```{r}
# Presentation of portfolio object
print(portf_maxret)
summary(portf_maxret)
```

***Optimazing Mean Return Portfolio***
```{r}
# Run the optimization --> Method = ROI
opt_maxret <- optimize.portfolio(R=stock.returns.monthly_xts, portfolio=portf_maxret,
                                 optimize_method="ROI", trace=TRUE)
print(opt_maxret)
summary(opt_maxret)
```

```{r}
# The opt_maxret object is of class optimize.portfolio.ROI and contains the following elements.
names(opt_maxret)
```

```{r}
# The value of the objective function at the optimum
extractStats(opt_maxret)
# The optimal weights at the optimum
extractWeights(opt_maxret)
```

```{r}
# The plot method charts of the optimal weights with the box constraints along with the optimal portfolio in risk-return space. The blue dots are the optimal weights and the gray triangles are the min and max of the box constraints.
plot(opt_maxret, 
     chart.assets=TRUE, 
     xlim=c(0.06, 0.25),
     ylim=c(-0.01, 0.01), 
     main="Maximizing Mean Return with ROI ")
```

```{r}
# The optimal portfolio can be plotted in "risk-return" space along with other feasible portfolios. The return metric is defined in the return.col argument and the risk metric is defined in the risk.col argument. The scatter chart includes the optimal portfolio (blue dot) and other feasible portfolios (gray circles) to show the overall feasible space given the constraints.
chart.RiskReward(opt_maxret,
                 eighbors = NULL,
                 return.col="mean", risk.col="sd",
                 element.color = "darkgray",
                 chart.assets=TRUE, 
                 xlim=c(0.02, 0.2), 
                 ylim = NULL, 
                 rp = FALSE,
                 main="Maximizing Mean Return with ROI")
```
```{r}
maxret.ef <- create.EfficientFrontier(R=stock.returns.monthly_xts, 
                                       portfolio=portf_maxret, 
                                       type="mean-StdDev")
chart.EfficientFrontier(maxret.ef, match.col="StdDev", type="l")
```

***Rebalancing Mean Return Portfolio***
An out of sample backtest is run with "optimize.portfolio.rebalancing". In this example, an initial training period of 36 months is used and the portfolio is rebalanced quarterly.
```{r}
# Backtesting with
# optimize_method   =   ROI
# rebalance_on      =   month
# training_period   =   2
bt_maxret <- optimize.portfolio.rebalancing(R=stock.returns.monthly_xts ,portfolio=portf_maxret,
                                            optimize_method="ROI",
                                            rebalance_on="month",
                                            training_period=20)
# short overview of the backtesting weights over time
head(extractWeights(bt_maxret)) 
# Weights over time 
chart.Weights(bt_maxret, main="Optimal Weights and Objective Measure at each Rebalance Period of MRP")
# summary performance 
returns_maxret <- Return.portfolio(R = stock.returns.monthly_xts, weights = extractWeights(bt_maxret))
returns_maxret[1:4,1]
charts.PerformanceSummary(returns_maxret)
charts.PerformanceSummary(cbind (returns_maxret, stock.returns.monthly_xts),
                          main="Performance summary with Portfolio MMR",
                          event.labels = TRUE,
                          wealth.index = TRUE)
```

### *b) Minimum-Variance*
***Minimizing Portfolio Variance***
```{r}
# Create portfolio object
portf_minvar <- portfolio.spec(assets=stockselection_1)
# Add full investment constraint to the portfolio object.The only constraint specified is the full investment constraint!
portf_minvar <- add.constraint(portfolio=portf_minvar, type="full_investment")
# Add objective to minimize variance
portf_minvar <- add.objective(portfolio=portf_minvar, type="risk", name="var")
```

***Optimazation Minimum Variance Portfolio***
```{r}
# Run the optimization
opt_gmv <- optimize.portfolio(R=stock.returns.monthly_xts, portfolio=portf_minvar,
                              optimize_method="ROI", trace=TRUE)
print(opt_gmv)
plot(opt_gmv, chart.assets=TRUE, xlim=c(0.06, 0.25),ylim=c(-0.01, 0.01), main="Minimizing Portfolio Varianc")
chart.RiskReward(opt_gmv,
                 neighbors = opt_maxret,
                 return.col="mean", risk.col="sd",
                 element.color = "darkgray",
                 chart.assets=TRUE, 
                 xlim=c(0.02, 0.2), 
                 ylim = NULL, 
                 rp = FALSE,
                 main="Minimizing Portfolio Varianc ")
```
```{r}
minvar.ef <- create.EfficientFrontier(R=stock.returns.monthly_xts, 
                                       portfolio=portf_minvar, 
                                       type="mean-StdDev")
chart.EfficientFrontier(minvar.ef, match.col="StdDev", type="l")
```

***Rebalancing Minimum Variance Portfolio***
```{r}
bt_gmv <- optimize.portfolio.rebalancing(R=stock.returns.monthly_xts,portfolio=portf_minvar,
                                         optimize_method="ROI",
                                         rebalance_on="month",
                                         training_period=20)
# short overview of the backtesting weights over time
head(extractWeights(bt_gmv)) 
# Weights over time 
chart.Weights(bt_gmv,
              main="Optimal Weights and Objective Measure at each Rebalance Period MVP.")
# summary performance 
returns_gmv <- Return.portfolio(R = stock.returns.monthly_xts, weights = extractWeights(bt_gmv))
returns_gmv[1:4,1]
charts.PerformanceSummary(returns_gmv)
p.returns<-merge(returns_gmv,returns_maxret)
names(p.returns)<-c("Rolling Global MaxRet","Rolling MaxRet")
charts.PerformanceSummary(p.returns,
                          main="Performance summary with Portfolio MaxRet and MinVar",
                          event.labels = TRUE,
                          wealth.index = TRUE)
```

***Constrained Minimum Variance Portfolio***
```{r}
# Add long only constraints
portf_minvar <- add.constraint(portfolio=portf_minvar, type="box",
                               min=0, max=1)
# Add group constraints
portf_minvar <- add.constraint(portfolio=portf_minvar,
                               type="group",
                               groups=list(groupA=1,
                                           groupB=c(2, 3),
                                           groupC=4),
                               group_min=c(0, 0.25, 0.10),
                               group_max=c(0.45, 0.6, 0.5))
```

***Optimization of Constrained Minimum Variance Portfolio***
```{r}
# Run the optimization
opt_minvar <- optimize.portfolio(R=stock.returns.monthly_xts, portfolio=portf_minvar,
                                 optimize_method="ROI", trace=TRUE)
print(opt_minvar)
extractStats(opt_minvar, prefix = NULL)
plot(opt_minvar, chart.assets=TRUE, xlim=c(0.06, 0.25),ylim=c(-0.01, 0.01), main="Constrained Minimum Variance Portfoli")
chart.RiskReward(opt_minvar,
                 neighbors = NULL,
                 return.col="mean", risk.col="sd",
                 element.color = "darkgray",
                 chart.assets=TRUE, 
                 xlim=c(0.02, 0.2), 
                 ylim = NULL, 
                 rp = FALSE,
                 main="Constrained Minimizing Portfolio Varianc ")
```


***Rebalancing of Constrained Minimum Variance Portfolio***
```{r}
bt_minvar <- optimize.portfolio.rebalancing(R=stock.returns.monthly_xts, portfolio=portf_minvar,
                                            optimize_method="ROI",
                                            rebalance_on="month",
                                            training_period=20)
bt_minvar
# short overview of the backtesting weights over time
head(extractWeights(bt_minvar)) 
# Weights over time 
chart.Weights(bt_minvar,
              main="Optimal Weights and Objective 
              Measure at each Rebalance Period of Constrained MVP.")
# summary performance 
returns_minvar <- Return.portfolio(R = stock.returns.monthly_xts, weights = extractWeights(bt_minvar))
returns_minvar[1:4,1]
charts.PerformanceSummary(returns_minvar)
p.returns<-merge(returns_gmv,returns_maxret,returns_minvar)
names(p.returns)<-c("Rolling Global MaxRet","Rolling MaxRet","Rolling MinVar")
charts.PerformanceSummary(p.returns,
                          main="Performance summary with Portfolio MaxRet and MinVar",
                          event.labels = TRUE,
                          wealth.index = TRUE)
```


### **c) Maximum Quadratic Utility Portfolios**
checking for a variety of constraints (constraints that can be computed with the `ROI`-solver) and different rebalancing periods (as well as rolling windows/training periods) to find, what might deliver you the best portfolios performance (use appropriate statistics to decide on that).
***Maximizing Quadratic Utility***
```{r}
# Create initial portfolio object
init_portf <- portfolio.spec(assets=stockselection_1)
# Create full investment constraint
fi_constr <- weight_sum_constraint(type="full_investment")
# Create long only constraint
lo_constr <- box_constraint(type="long_only", assets=init_portf$assets)
# Combine the constraints in a list
qu_constr <- list(fi_constr, lo_constr)
# Create return objective
ret_obj <- return_objective(name="mean")
# Create variance objective specifying a risk_aversion parameter which controls
# how much the variance is penalized
var_obj <- portfolio_risk_objective(name="var", risk_aversion=0.25)
# Combine the objectives into a list
qu_obj <- list(ret_obj, var_obj)
```

***Optimazation Maximizing Quadratic Utility***
```{r}
# Run the optimization
opt_qu <- optimize.portfolio(R=stock.returns.monthly_xts, portfolio=init_portf,
                             constraints=qu_constr,
                             objectives=qu_obj,
                             optimize_method="ROI",
                             trace=TRUE)
opt_qu
plot(opt_qu, chart.assets=TRUE, xlim=c(0.06, 0.25),ylim=c(-0.01, 0.01), main="Maximizing Quadratic Utility")
chart.RiskReward(opt_qu,
                 neighbors = NULL,
                 return.col="mean", risk.col="sd",
                 element.color = "darkgray",
                 chart.assets=TRUE, 
                 xlim=c(0.02, 0.2), 
                 ylim = NULL, 
                 rp = FALSE,
                 main="Maximizing Quadratic Utility ")
```

***Rebalancing Maximizing Quadratic Utility***
```{r}
bt_qu <- optimize.portfolio.rebalancing(R=stock.returns.monthly_xts, portfolio=init_portf,
                                        constraints=qu_constr,
                                        objectives=qu_obj,
                                        optimize_method="ROI",
                                        rebalance_on="quarter",
                                        training_period=36)
# short overview of the backtesting weights over time
head(extractWeights(bt_qu)) 
# Weights over time 
chart.Weights(bt_qu,
              main="Optimal Weights and Objective Measure at each Rebalance Period of MQU")
# summary performance 
returns_qu <- Return.portfolio(R = stock.returns.monthly_xts, weights = extractWeights(bt_qu))
returns_qu[1:4,1]
charts.PerformanceSummary(returns_qu)
p.returns<-merge(returns_gmv,returns_maxret,returns_minvar, returns_qu)
names(p.returns)<-c("Rolling Global MaxRet","Rolling MaxRet","Rolling MinVar","Rolling Quadratic Utility")
charts.PerformanceSummary(p.returns,
                          main="Performance summary with Portfolio MaxRet and MinVar and QuadUti",
                          event.labels = TRUE,
                          wealth.index = TRUE)
```

```{r}
# Everything together ;)
aa <- cbind (p.returns, stock.returns.monthly_xts)
charts.PerformanceSummary(aa,
                          main="Performance summary with Portfolio MaxRet and MinVar and QuadUti",
                          event.labels = TRUE,
                          wealth.index = TRUE)
```



## Exercise 2: Custom moments function

Check `vignette("custom_moments_objectives")` to implement a variety of robust covariance matrix estimates (see `?MASS::cov.rob`, `?PerformanceAnalytics::ShrinkageMoments` and maybe `?PerformanceAnalytics::EWMAMoments` - the latter one only for backtesting) for the minimum variance and quadratic utility portfolios. Plot the different Efficient frontiers, optimal portfolios and weights and visualize the different covariances. Also make yourselves comfortable with cleaning outliers from your timeseroies via `return.Clean()`.

Check `vignette("custom_moments_objectives")` to implement a variety of robust covariance matrix estimates (see `?MASS::cov.rob`, `?PerformanceAnalytics::ShrinkageMoments` and maybe `?PerformanceAnalytics::EWMAMoments` - the latter one only for backtesting) for the minimum variance and quadratic utility portfolios. Plot the different Efficient frontiers, optimal portfolios and weights and visualize the different covariances. Also make yourselves comfortable with cleaning outliers from your timeseries via `return.Clean()`.


```{r}
require(timetk)
stockselection_2 <- c("AAPL", "MSFT", "AMZN", "PFE", "JPM", "CSCO", "XOM", "BAC", "NKE", "WMT")
n_2 <- length(stockselection_2)
stock.prices <- stockselection_2 %>%
  tq_get(get  = "stock.prices", from = "2000-01-01",to = Sys.Date( )) %>%
  group_by(symbol)
stock.prices %>% slice(1:2) 
stock.returns.monthly <- stock.prices %>%  
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns"
               )
stock.returns.monthly %>% slice(1:2) 
#Make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts
stock.returns.monthly_xts_withoutliners <- pivot_wider(stock.returns.monthly,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>% 
  tk_xts(date_var = date, silent = TRUE)
stock.returns.monthly_xts <- Return.clean(stock.returns.monthly_xts_withoutliners, method = "boudt", alpha = 0.01)
```

**Now, we create the initial minimum variance portfolio**

In a first step, we require the necessary packages.
Then, we construct the initial portfolio with basic constraints. We construct the portfolio in a way that R minimizes the standard deviation.

Usually we want to invest our entire budget and therefore set type="full_investment" which sets the sum of weights to 1. Alternatively we can set the type="weight_sum" to have mimimum/maximum weight_sum equal to 1.

```{r}
require(PortfolioAnalytics)
require(DEoptim)
require(ROI)
require(ROI.plugin.glpk)
require(ROI.plugin.quadprog)
# Construct initial portfolio with basic constraints.
init.port.minv <- portfolio.spec(assets=colnames(stock.returns.monthly_xts),category_labels = stockselection_2)
init.port.minv <- add.constraint(portfolio=init.port.minv, type="full_investment")
init.port.minv <- add.constraint(portfolio=init.port.minv, type="long_only")
#Portfolio with standard deviation as an objective
SD.port.minv <- add.objective(portfolio=init.port.minv, type="risk", name="StdDev")
```

**Next, we create initial maximum quadratic utility portfolio. **
We construct the initial quadratic utility portfolio with the basic constraints ( fullinvestment, long_only). 
```{r}
# Construct initial portfolio with basic constraints.
init.port.maxq <- portfolio.spec(assets=colnames(stock.returns.monthly_xts),category_labels = stockselection_2)
#init.port.maxq <- add.constraint(init.port.maxq, type = "box", min = 0, max = 1)
init.port.maxq <- add.constraint(portfolio=init.port.maxq, type="full_investment")
init.port.maxq <- add.constraint(portfolio=init.port.maxq, type="long_only")
#Portfolio with standard deviation as an objective
SD.port.maxq <- add.objective(portfolio=init.port.maxq, type="return", name="mean")
SD.port.maxq <- add.objective(portfolio=SD.port.maxq, type="risk", name="var", risk_aversion=0.25)
SD.port.maxq <- add.objective(portfolio=init.port.maxq, type="quadratic_utility", risk_aversion=0.25)
SD.port.maxq
```


**function to estimate covariance matrix with cov.rob for minimum variance**
Description of cov.rob

Compute a multivariate location and scale estimate with a high breakdown point – this can be thought of as estimating the mean and covariance of the good part of the data. cov.mve and cov.mcd are compatibility wrappers.

In method "mcd" it is the volume of the Gaussian confidence ellipsoid, equivalently the determinant of the classical covariance matrix, that is minimized. The mean of the subset provides a first estimate of the location, and the rescaled covariance matrix a first estimate of scatter. The Mahalanobis distances of all the points from the location estimate for this covariance matrix are calculated, and those points within the 97.5% point under Gaussian assumptions are declared to be good. The final estimates are the mean and rescaled covariance of the good points.

```{r}
sigma.robust <- function(R){
    require(MASS)
    out <- list()
    out$sigmarob <- cov.rob(R, method="mcd")$cov
    return(out)
}
sigmarob <- sigma.robust(stock.returns.monthly_xts)$sigmarob
sigmarob
```



**function to estimate covariance matrix with ShrinkageMoments for minimum variance** 
Definition of Shrinkeage: Shrinkage is where extreme values in a sample are “shrunk” towards a central value, like the sample mean.

```{r}
sigma.robust.shrink <- function(R){
    targets <- c(1,3,4)
    f <- rowSums(stock.returns.monthly_xts)
    out <- list()
    out$sigmashrink <- M2.shrink(stock.returns.monthly_xts, targets, f)$M2sh
    return(out)
}
sigma.shrink <- sigma.robust.shrink(stock.returns.monthly_xts)$sigmashrink
sigma.shrink
```

**Optimize portfolios**
Now we can use the custom moment function in optimize.portfolio to estimate the solution
to the minimum standard deviation portfolio.

Here we extract the weights and compute the portfolio standard deviation to verify that the
the robust estimate of the covariance matrix was used in the optimization.

```{r message=FALSE, warning=FALSE}
#portfolio moment "cov.rob"
##minimum variance portfolio
opt.sd.minv <- optimize.portfolio(stock.returns.monthly_xts, SD.port.minv, optimize_method="ROI", momentFUN="sigma.robust", trace = TRUE)
##maximum quadratic utility portfolio
opt.sd.maxq <- optimize.portfolio(stock.returns.monthly_xts, SD.port.maxq, optimize_method="ROI", momentFUN="sigma.robust", trace = TRUE)
#portfolio moment "ShrinkeageMoments"
##minimum variance portfolio
opt.sd.minv.shrink <- optimize.portfolio(stock.returns.monthly_xts, SD.port.minv, optimize_method="ROI", momentFUN="sigma.robust.shrink", trace = TRUE)
##maximum quadratic utility portfolio
opt.sd.maxq.shrink <- optimize.portfolio(R=stock.returns.monthly_xts, portfolio=SD.port.maxq, optimize_method="ROI", momentFUN="sigma.robust.shrink", trace = TRUE)
weights <- extractWeights(opt.sd.minv)
sigmarob <- sigma.robust(stock.returns.monthly_xts)$sigmarob
sqrt(t(weights) %*% sigmarob %*% weights)
extractObjectiveMeasures(opt.sd.minv)$StdDev
opt.sd.minv
```
**Plot the covariance matrix from cov.rob**
```{r echo = FALSE}
chart.Correlation(sigmarob, histogram = TRUE)
```

**Plot the covariance matrix from shrink**
```{r echo = FALSE}
chart.Correlation(sigma.shrink, histogram = TRUE)
```
**Plot the Portfolios**
```{r}
plot(opt.sd.minv, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02))
plot(opt.sd.minv.shrink, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02))
plot(opt.sd.maxq, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.05))
plot(opt.sd.maxq.shrink, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.05))
```
**Chart Efficient Frontiert for the minimum variance Portfolio**
```{r echo = FALSE}
prt_eff_minv <- create.EfficientFrontier(R=stock.returns.monthly_xts, portfolio=SD.port.minv, type="mean-StdDev", match.col = "StdDev")
chart.EfficientFrontier(prt_eff_minv, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)
chart.EF.Weights(prt_eff_minv, colorset=rainbow(n = length(stockselection_2)), match.col="StdDev", cex.lab = 1, main = "StdDev")
```
**Chart Efficient Frontiert for the quadratic utility Portfolio**
```{r echo = FALSE}
prt_eff_maxq <- create.EfficientFrontier(R=stock.returns.monthly_xts, portfolio=SD.port.maxq, type="mean-StdDev", match.col = "StdDev")
chart.EfficientFrontier(prt_eff_maxq, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)
chart.EF.Weights(prt_eff_maxq, colorset=rainbow(n = length(stockselection_2)), match.col="StdDev", cex.lab = 1, main = "StdDev")
```


## Exercise 3: Regime Switching

Have a look at `demo(regime_switching)` and estimate and rebalance portfolios based on 2/3 regimes. Can you plot the regimes over time?


```{r}
demo(regime_switching)
```


```{r}
#' Load package and data.
library(PortfolioAnalytics)
```


```{r}
# get monthly stock returns from the S&P500
monthly_returnsSP500 <- "^GSPC" %>%
  tq_get(get = "stock.prices", from = "2000-01-01", to = "2019-08-31") %>%
  tq_transmute(adjusted, periodReturn, period = "monthly", col_rename = "returns SP500")
monthly_returnsSP500
```

```{r}
#calculate the rolling mean monthly
rollmeanSP500 <- rollmean(monthly_returnsSP500[, "returns SP500"], 2)
rollmeanSP500
```

```{r}
vector <- c(rollmeanSP500)
#2=good economy, 1=bad economy
regime1or2 <-as.numeric(vector>0)+1
regime1or2
```


```{r}
SP500dates <- select(monthly_returnsSP500,date)
#regime 1 is bad market phase (1) and regime 2 is good market phase (2)
data_frame <- data.frame(SP500dates[2:236,], regime1or2)
data_frame
#convert it in xts format
regime_xts <- data_frame %>%
  select(date,regime1or2) %>%
  tk_xts(silent = TRUE)
regime_xts
```



```{r}
stockselection <- c("MO", "MDLZ", "PEP", "PM", "KHC", "T","WMT", "MSFT","BAC","PG", "^GSPC")
#Get the prices of the stocks to transmute it to returns
stock.prices <- stockselection %>%
  tq_get(get  = "stock.prices", from = "2000-01-01",to = "2018-08-31") %>%
  group_by(symbol)
#Create monthly returns
stock.returns.monthly <- stock.prices %>%  
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns"
               )
#make a tibble with dates and returns for all stocks
#Make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts (necessary for Portfolioanalytics)
R <- pivot_wider(stock.returns.monthly,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>% 
  tk_xts(date_var = date, silent = TRUE)
colnames(R) <- c("AAPL", "MSFT", "AMZN", "JPM", "AMGN", "CSCO", "GILD", "MCD", "QCOM", "WMT")
funds <- colnames(R)
#Amgen -bio technology
#Gild - pharmacy
#QCOM - tele communication 
R %>% head()
```

#########ignore: old solution###########
```{r}
#' Here I just randomly samples values to create regime 1 or regime 2. In 
#' practice, this could based on volatility or other regime switching models
# 
set.seed(123)
regime <- xts(sample(1:2, nrow(R), replace=TRUE, prob=c(0.3, 0.7)), index(R))
#Regimes are simply market periods with different parameters, such as good market phases (upswing) and bad (downturn). Regimes are different economic environments (e.g. good times and bad times). 
# We want  a different portfolio for good times and for bad times (and maybe - as a third regime - for times that are uncertain but neither really good nor bad). --> 
#here, they assign different dates to regime 1 or 2
#there are more in regime 2 because prob. of regime 1 is 30% and of regime 2 is 70%
regime
```
##################################################

```{r}
#' Construct portfolio for regime 1 - bad economy.
#here, the first regime is considered with a risk approach (Mean-ES portfolio and other constraints)   --> we optimize ES
#Es = Conditional Value at risk: considers losses that exceed the value-at-risk and determines their average amount. 
port1 <- portfolio.spec(funds)
port1 <- add.constraint(port1, "weight_sum", min_sum=0.99, max_sum=1.01)
port1 <- add.constraint(port1, "box", min=0.05, max=0.5)
port1 <- add.objective(port1, type="risk", name="ES", arguments=list(p=0.9))
port1 <- add.objective(port1, type="risk_budget", name="ES", 
                       arguments=list(p=0.9), max_prisk=0.5)
```

```{r}
#' Construct portfolio for regime 2 - good economy.
#here regime 2 is a regime based on standard investment with volatility - here we used the standard deviation --> we optimize Stdev
port2 <- portfolio.spec(funds)
port2 <- add.constraint(port2, "weight_sum", min_sum=0.99, max_sum=1.01)
port2 <- add.constraint(port2, "box", min=0, max=0.6)
port2 <- add.objective(port2, type="risk", name="StdDev")
port2 <- add.objective(port2, type="risk_budget", name="StdDev", max_prisk=0.5)
```

```{r}
#' Combine the portfolios.
portfolios <- combine.portfolios(list(port1, port2))
#' Now we construct the regime model and corresponding portfolios to use for
#' each regime.
# we merge the portfolios and the regimes (becauase we cannot merge every single portfolio with the regimes)
regime.port <- regime.portfolios(regime_xts, portfolios)
regime.port
```



```{r}
#' This optimization should result in out portfolio for regime 2 (optimizing Stdev).
opt1 <- optimize.portfolio(R, regime.port, 
                           optimize_method="random", 
                           search_size=2000, 
                           trace=TRUE)
opt1
opt1$regime
```

```{r}
#' This optimization should result in out portfolio for regime 1.
opt2 <- optimize.portfolio(R[1:(nrow(R)-1)], regime.port, 
                           optimize_method="DEoptim", 
                           search_size=2000, 
                           trace=TRUE)
```

```{r}
opt2
opt2$regime
```


```{r}
#' Run optimization with rebalancing using our regime switching portfolio.
opt.rebal <- optimize.portfolio.rebalancing(R, regime.port,
                                            optimize_method="random", 
                                            rebalance_on="quarters", 
                                            training_period=130,
                                            search_size=2000, 
                                            trace=TRUE)
#' The print and summary methods work the same as they do for optimizations 
#' without regime switching.
opt.rebal
summary(opt.rebal)
```


```{r}
#' We can extract which regime portfolio we optimized with at each rebalance date.  
#if there are structural changes in the data series, maybe a date fits better in the other regime now
lapply(opt.rebal$opt_rebalancing, function(x) x$regime)
```


```{r}
#' Extract the optimal weights at each rebalance date.
wt <- extractWeights(opt.rebal)
wt
```


```{r}
#' Extract the objective measures*.
obj <- extractObjectiveMeasures(opt.rebal)
str(obj)
obj
```


```{r}
# Extract the stats*.
xt <- extractStats(opt.rebal)
str(xt)
```


```{r}
#' *
#' Note that this returns a list of N elements for N regimes. We may have 
#' different objectives and/or a different number of objectives which makes
#' returning a single xts object difficult/
#' Extract the optimal weights at each rebalance date.
chart.Weights(opt.rebal, colorset=rainbow10equal)
wt
```

```{r}
#' Chart the risk contribution for regime 1
# 
chart.RiskBudget(opt.rebal, match.col="ES", risk.type="percentage", 
                 regime=1, colorset=rainbow10equal)
opt2
```

```{r}
#' Chart the risk contribution for regime 2
chart.RiskBudget(opt.rebal, match.col="StdDev", risk.type="percentage", 
                 regime=2, colorset=rainbow10equal)
opt1
```


## Exercise 4: Single Index-Model*

*Now we are going to estimate the Portfolio Input Parameters with the Single-Index Model. Use your ten assets and additionally choose the S&P500 as index (same returns etc).*

*a) Regress all stocks on the index. Show alpha, beta and residual variance. Calculate systematic and firm-specific risk. Are there any significant alphas? (You should double check with the appropriate `PerformanceAnalytics` Functions)*

*b) Extract the betas and calculate systematic and unsystematic risk, derive the whole covariance matrix. To do this you can use _CH15_Factor_Modfels_for_Asset_Returns.pdf (15.3.1)_ and the code implemented in the function sharpeFactorEstimator that you find [here](http://financewithr.blogspot.com/2013/06/portfolio-optimization-using-single.html) (please do not just copy everything, but try to understand what you are doing, e.g. check why and if G.hat has the same values as found by the multivariate regression).*

*c) Now use the _custom-moments_ functions from Exercise 2 to implement the single-factor model into the portfolio optimization framework and plot the efficient frontier using the parameters estimated by the single factor model next to the EF of the full-covariance model. Calculate MVP, TP etc. and work out the differences in weights, portfolio return and portfolio risk.*


### Get the packages 

```{r echo=FALSE}
install.packages("fEcofin", repos="http://R-Forge.R-project.org")
library(fEcofin)
library(fPortfolio)
```


Now we are going to estimate the Portfolio Input Parameters with the Single-Index Model. Use your ten assets and additionally choose the S&P500 as index (same returns etc).

```{r}
stockselection_4 <- c("MO", "MDLZ", "PEP", "PM", "KHC", "T","WMT", "MSFT","BAC","PG", "^GSPC")
stockselection_4
# Presettings
n <- length(stockselection_4)
#Get the prices of the stocks
stock.prices_4 <- stockselection_4 %>%
  tq_get(get  = "stock.prices", from = "2015-07-06",to = Sys.Date( )) %>% #First trade day of KHC
  group_by(symbol)
# Output the first two entries of each stock!
stock.prices_4 %>% slice(1:2) 
stock.prices_4 %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  ggtitle("Price chart for all stocks - all in one")
# Plotting the stock prices in each frame
stock.prices_4 %>%
  ggplot(aes(x = date, y = adjusted)) +
  geom_line() +
  facet_wrap(~symbol, scales = "free_y") +
  theme_classic() +
  labs(x = "Date", y = "Price") +
  ggtitle("Price chart all stocks - in each frame") 
```

```{r}
# Create monthly returns by the tq_transmute() = adds new variables to an existing tibble;
stock.returns.monthly_4 <- stock.prices_4 %>%  
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns")
# Output the first two entries of each stock!
stock.returns.monthly_4 %>% slice(1:2) 
# Make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts
stock.returns.monthly_xts_4 <- pivot_wider(stock.returns.monthly_4,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>%
   tk_xts(date_var = date, silent = TRUE)
# Output the first entries (simple returns from adjusted prices) of each stock!
stock.returns.monthly_xts_4[1]
# Plotting a performance summary chart 
charts.PerformanceSummary(stock.returns.monthly_xts_4, 
                          main="Performance summary")
```

a) Regress all stocks on the index. Show alpha, beta and residual variance. Calculate systematic and firm-specific risk. Are there any significant alphas? (You should double check with the appropriate `PerformanceAnalytics` Functions)

```{r}
#Regress all stocks on the index
alpha.Stocks <- CAPM.alpha(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], Rf = 0)
beta.Stocks <- CAPM.beta(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], Rf = 0)
StdDev.Index <- StdDev(R = stock.returns.monthly_xts_4[,n],
                   clean = "none",
                   method = "pearson")
lm(stock.returns.monthly_xts_4[,-n] ~ stock.returns.monthly_xts_4[,n])
for(i in 1:n) {
plot.default(x = stock.returns.monthly_xts_4[, n], 
            y = stock.returns.monthly_xts_4[, i], 
            main = stockselection_4[i], 
            xlab = "Index Returns", 
            ylab = "Stock Returns", 
            abline(lm(stock.returns.monthly_xts_4[, i] ~ stock.returns.monthly_xts_4[, n])))
}
#Calculate systematic (Market-Specific) Risk by mulitplying Variance (StdDev^2) of the S&P500 and the Beta^2 of each stock
sys.risk <- SystematicRisk(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], Rf = 0)
sys.risk
#Calculate Firm-specific Risk / Residual Variance
firm.specific.risk <- SpecificRisk(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], Rf = 0)
firm.specific.risk
#Summary
summary.SFM <- table.SFM(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], scale = NA, Rf = 0, digits = 6)
summary.SFM
  
# Tracking Error: Der Nachbildungsfehler[1] (englisch tracking error, Abkürzung TE) bezeichnet die ungewollte Abweichung zwischen der Wertentwicklung eines Indexfonds oder eines Portfolios gegenüber der seiner Benchmark über einen bestimmten Beobachtungszeitraum.[2]
```

b) Extract the betas and calculate systematic and unsystematic risk, derive the whole covariance matrix. To do this you can use _CH15_Factor_Models_for_Asset_Returns.pdf (15.3.1)_ and the code
implemented in the function sharpeFactorEstimator that you find [here](http://financewithr.blogspot.com/2013/06/portfolio-optimization-using-single.html) (please do not just copy everything, but try to understand what you are doing, e.g. check why and if G.hat has the same values as found by the multivariate regression).

```{r}
#Calculate Beta of Portfolio by average each stocks beta
beta.portfolio <- mean(beta.Stocks)
beta.portfolio
#Calculate systematic (Market-Specific) Risk of portfolio
sys.risk.portfolio <- mean(sys.risk)
sys.risk.portfolio
#Calculate unsystematic risk by calculating the mean of the firm-specific risk
unsys.risk.portfolio <- mean(firm.specific.risk)
unsys.risk.portfolio
#Calculate Covariance Matrix
stock.returns.monthly_data <- as.data.frame((stock.returns.monthly_xts_4))
head(stock.returns.monthly_data)
returns <- as.timeSeries(stock.returns.monthly_data[,-n])
factors <- as.vector(as.timeSeries(stock.returns.monthly_data)[,n])
names(data)
data <- returns
attr(data, "factors") <- factors
nScenarios <- nrow(data)
X.mat <- cbind(rep(1, times=nScenarios), factors)
G.hat <- solve(qr(X.mat), data) #G.hat is alpha
beta.hat <- G.hat[2, ] #is beta
eps.hat <- data - X.mat %*% G.hat
diagD.hat <- diag(crossprod(eps.hat) / (nScenarios-2))
cov.si = var(factors)*(beta.hat%o%beta.hat) + diag(diagD.hat)
cov.si
```

c) Now use the _custom-moments_ functions from Exercise 2 to implement the single-factor model into the portfolo optimization framework and plot the efficient frontier using the parameters estimated by the single factor model next to the EF of the full-covariance model.

```{r}
#Function to implement single-factor model into the portfolio optimization framework
data(stock.returns.monthly_xts_4)
returns <- as.timeSeries(stock.returns.monthly_xts_4)
names(data)
data <- returns[, -c(n)]
factors <- returns[, n]
attr(data, "factors") <- factors
# Sharpe's Single Index Factor Model:
sharpeFactorEstimator <- 
function(x, spec=NULL, ...)
{
    # Sharpe Single Index Model:
    data <- getDataPart(x)
    factors <- attr(x, "factors")
    nScenarios <- nrow(data)
    X.mat <- cbind(rep(1, times=nScenarios), factors)
    G.hat <- solve(qr(X.mat), data)
    beta.hat <- G.hat[2, ]
    eps.hat <- data - X.mat %*% G.hat
    diagD.hat <- diag(crossprod(eps.hat) / (nScenarios-2))
    mu <- G.hat[1, ] + G.hat[2, ] * colMeans(factors)  
    Sigma <- var(factors)[[1]] * (beta.hat %o% beta.hat) + diag(diagD.hat)
    
    # Return Value:
    list(mu = mu, Sigma = Sigma)
}
 #Chart the efficient frontier using the parameters estimated by the single factor model
sharpe_1 <- portfolioFrontier(data)
tailoredFrontierPlot(sharpe_1)
points(frontierPoints(sharpe), col = "steelblue")
#Chart Efficient Frontier minimum variance
prt_eff_minv <- create.EfficientFrontier(R=stock.returns.monthly_xts_4, portfolio=SD.port.minv, type="mean-StdDev", match.col = "StdDev")
chart.EfficientFrontier(prt_eff_minv, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)
```

Calculate MVP, TP etc. and work out the differences in weights, portfolio return and portfolio risk.

```{r}
#weights, portfolio return and portfolio risk of MVP
opt.sd.minv.shrink
plot(opt.sd.minv.shrink, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02)) 
#weights, portfolio return and portfolio risk of QUP
opt.sd.maxq.shrink
plot(opt.sd.maxq.shrink, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(-0.08,0.05))
#weights, portfolio return and portfolio risk of tangency portfolio (Highest risk/return ratio)
weight_tp_sharpe <- sharpe_1@portfolio@portfolio[["weights"]][26, ]
return_tp_sharpe <- sharpe_1@portfolio@portfolio[["targetReturn"]][26, ]
risk_tp_sharpe <- sharpe_1@portfolio@portfolio[["targetRisk"]][26, ]
weight_tp_sharpe
return_tp_sharpe
risk_tp_sharpe
````




















