---
title: "Portfoliomanagement and Financial Analysis - Assignment 4"
subtitle: "Submit until Monday 2019-10-07, 13:00"
author: "Can, Merve"
output: html_notebook
---

```{r load_packs}
pacman::p_load(tidyverse,tidyquant,PortfolioAnalytics,nloptr)
```

**Please** remember to put your assignment solutions in `rmd` format using **many** chunks and putting readable text in between, similar to my examples given in Research Methods and Assignment 1! Also, each student has to select his own set of 10 stocks having data available as of `2000-01-01`. Select by Sharpe-ratio, dominance or any other method (e.g. matching your first name).

For all exercises: Please use the Assignment-Forum to post your questions, I will try my best to help you along!

These are the stocks that I will be working with

```{r}
library(timetk)

SP500 <- tq_index("SP500")
NASDAQ <- tq_exchange("NASDAQ")
NYSE <- tq_exchange("NYSE") 

stocks.selection <- SP500 %>% 
  inner_join(rbind(NYSE,NASDAQ) %>% select(symbol,last.sale.price,market.cap,ipo.year),by=c("symbol")) %>%
  filter(ipo.year<2000&!is.na(market.cap)) %>% 
  arrange(desc(weight)) %>% 
  slice(1:10)

stocks.selection
```

These are the returns of the selected stocks.

```{r}
stocks.returns <- stocks.selection$symbol %>%
    tq_get(get  = "stock.prices",
           from = "2000-01-01",
           to   = "2019-12-31") %>%
    group_by(symbol) %>%
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly")

stocks.returns

```
These are the stocks return in the xts format and also in a wide format
````{r}
stocks.returns.xts <- stocks.returns%>%
                      subset( select = c(symbol,date, monthly.returns)) %>%
                      pivot_wider(names_from = symbol, 
                                  values_from = monthly.returns) %>% 
                      tk_xts(date_var = date, silent = TRUE)

```


## Exercise 1: Rebalancing

Have a look at `vignette("ROI_vignette")` and the `optimize.portfolio.rebalancing` command. Use your dataset to compute 


### a) Maximizing Mean-Return (tangency portfolio)

First we create portfolio, add constraints and add an objective.
```{r}
portf_maxret <- portfolio.spec(assets=stocks.selection$symbol)%>%
  add.constraint(type="full_investment") %>%
  add.constraint(type="long_only") %>%
  add.objective(type="return", name="mean")

portf_maxret
```



Now we run the optimization.
```{r}
opt_maxret <- optimize.portfolio(R=stocks.returns.xts, portfolio=portf_maxret,
                                 optimize_method="ROI", trace=TRUE)
opt_maxret
```
Here we can see that we invest everythin into one stock, which is the NVIDIA Corporation.

Now we plot the mu-sigma diagram and the weights.
```{r}
 plot(opt_maxret, chart.assets=TRUE, main="Maximum Return", 
      xlim=c(0.05, 0.35), ylim=c(0,0.04))

```


Now we plot the efficient frontier.
```{r}
maxret.ef <- create.EfficientFrontier(R=stocks.returns.xts, 
                                       portfolio=portf_maxret, 
                                       type="mean-StdDev")

chart.EfficientFrontier(maxret.ef, match.col="StdDev", type="l")
```

Backtesting:  Quarterly rebalancing with 5 year training period
```{r}
bt_maxret <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_maxret,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period=60)
bt_maxret 
```

```{r}
chart.Weights(bt_maxret, main = "Rebalanced Weights Over Time", 
              colorset=rainbow(n = length(stocks.selection$symbol)))
```

Backtesting:  Quarterly rebalancing with 5 year training period and 4 year rolling window
```{r}
bt_maxret2 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_maxret,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period= 60,
                                            rolling_window = 48)
bt_maxret2
```


```{r}
chart.Weights(bt_maxret2, main = "Rebalanced Weights Over Time",
              colorset=rainbow(n = length(stocks.selection$symbol)))
```              

### b) Minimum-Variance

First we create a portfolio, add constraints and add an objective.
```{r}
portf_minvar <- portfolio.spec(assets= stocks.selection$symbol) %>%
  add.constraint(type="full_investment") %>%
  add.constraint(type = "long_only") %>%
  add.objective(type="risk", name="var")
  
portf_minvar
```

Now we run the optimization
```{r}
opt_minvar <- optimize.portfolio(R = stocks.returns.xts,portfolio = portf_minvar,
                              optimize_method = "ROI", trace = TRUE)

opt_minvar
```
Here we can see the optimal weights. We do not invest in Amazon, NVIDIA Corporation and Cisco Systems.

Now we plot the weights.
```{r}
plot(opt_minvar, chart.assets=TRUE, main="Minimum Variance with Constraints", 
      xlim=c(0.05, 0.35), ylim=c(0,0.05))
```



Backtesting: Quarterly rebalancing with 5 year training period
```{r}
bt_minvar <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_minvar,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period=60)
bt_minvar
```


```{r}
chart.Weights(bt_minvar, main = "Rebalanced Weights Over Time", 
              colorset=rainbow(n = length(stocks.selection$symbol)))
```


Backtesting: Quarterly rebalancing with 5 year training period and 4 year rolling window
```{r}
bt_minvar2 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_minvar,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period=60,
                                            rolling_window = 48)
bt_minvar2
```


```{r}
chart.Weights(bt_minvar2, main = "Rebalanced Weights Over Time", colorset=rainbow(n = length(stocks.selection$symbol)))
```

### c) Maximum Quadratic Utility Portfolios

First we create the portfolio and add constraints.
```{r}
portf_maxqua <- portfolio.spec(assets= stocks.selection$symbol) %>%
  add.constraint(type = "full_investment") %>%
  add.constraint(type = "long_only")
portf_maxqua
```
We will compute two quadratic utility portfolios, one with a risk aversion of 1 (not very risk averse investor) and one with a risk aversion of 5 (risk averse investor).

### Risk aversion = 1

Now we create return objective, risk aversion objective and combine.
```{r}
ret_obj <- return_objective(name="mean")

var_obj <- portfolio_risk_objective(name="var", risk_aversion=1)

qu_obj <- list(ret_obj, var_obj)

qu_obj
```

Now we run the optimization.
```{r}
opt_qu <- optimize.portfolio(R=stocks.returns.xts, portfolio=portf_maxqua,
                             objectives=qu_obj,
                             optimize_method="ROI",
                             trace=TRUE)
opt_qu
```
Here we can see that we invest into 5 securietes: Apple, Amazon, NVIDIA, Adobe and Amgen.

Now we plot the weights.
```{r}
plot(opt_qu, chart.assets=TRUE, main="Maximum Quadratic Utility", 
      xlim=c(0, 0.35), ylim=c(0,0.05))
```

Backtesting: Quarterly rebalancing with 5 year training period
```{r}
bt_qu <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60)
bt_qu
```

```{r}
chart.Weights(bt_qu, main = "Rebalanced Weights Over Time", 
              colorset=rainbow(n = length(stocks.selection$symbol)))
```


### Risk aversion = 5
Create risk aversion objective and combine with return objective
```{r}
var_obj2 <- portfolio_risk_objective(name="var", risk_aversion=5)

qu_obj2 <- list(ret_obj, var_obj2)

qu_obj2
```

Now run the optimization
```{r}
opt_qu2 <- optimize.portfolio(R=stocks.returns.xts, portfolio=portf_maxqua,
                             objectives=qu_obj2,
                             optimize_method="ROI",
                             trace=TRUE)
opt_qu2
```
Here we can see that we invest in everything except for Cisco, Qualcomm and Oracle.

Now we plot the weights.
```{r}
plot(opt_qu2, chart.assets=TRUE, main="Maximum Quadratic Utility", 
      xlim=c(0, 0.35), ylim=c(0,0.05))
```

Backtesting: Quarterly rebalancing with 5 year training period
```{r}
bt_qu3 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj2,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60)
bt_qu
```

```{r}
chart.Weights(bt_qu3, main = "Rebalanced Weights Over Time",
              colorset=rainbow(n = length(stocks.selection$symbol)))
```

Backtesting: Quarterly rebalancing with 5 year training period
```{r}
bt_qu4 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj2,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60,
                                        rolling_window = 48)
bt_qu4
```

```{r}
chart.Weights(bt_qu4, main = "Rebalanced Weights Over Time", 
              colorset=rainbow(n = length(stocks.selection$symbol)))
```

## Comparing Portfolio Performance

### Sigma and Mu of Portfolios
Now we are going to make a simple mu-sigma plot of the portfolios to see if there is a dominated portfolio.
```{r}
sigma <- c(StdDev(portfolio.returns.maxret$Ra),
           StdDev(portfolio.returns.minvar$Ra),
           StdDev(portfolio.returns.qu$Ra),
           StdDev(portfolio.returns.qu2$Ra))

mu <- c(mean(portfolio.returns.maxret$Ra),
        mean(portfolio.returns.minvar$Ra),
        mean(portfolio.returns.qu$Ra),
        mean(portfolio.returns.qu2$Ra))

sigma.mu <- cbind(sigma,mu)

sigma.mu
```

```{r}
plot(sigma.mu)
```
As we can see here, there is no portfolio that is dominated by another portfolio.
weights of the portfolios. Since we could not decide for a portfolio with the mu-sigma plot, we will compare all portfolios to a benchmark, in this case the S&P 500 index because all the stocks are part of this index, and try to evaluate the performances of each portfolio with the CAPM.

### CAPM
The first step is to extract the optimal weight of the portfolios into datasets.
```{r}
wts_maxret <- extractWeights(opt_maxret)
wts_minvar <- extractWeights(opt_minvar)
wts_qu <- extractWeights(opt_qu)
wts_qu2 <- extractWeights(opt_qu2)

wts_maxret
wts_minvar
wts_qu 
wts_qu2
```

Now we will calculate the returns of each porfolio.
```{r}
stocks.returns

portfolio.returns.maxret <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_maxret, 
                 col_rename  = "Ra")

portfolio.returns.maxret

portfolio.returns.minvar <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_minvar, 
                 col_rename  = "Ra")

portfolio.returns.minvar

portfolio.returns.qu <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_qu, 
                 col_rename  = "Ra")

portfolio.returns.qu

portfolio.returns.qu2 <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_qu2, 
                 col_rename  = "Ra")

portfolio.returns.qu2
```

In this step we will calculate the baseline S&P 500 returns.
```{r}
baseline.returns <- "^GSPC" %>%
    tq_get(get  = "stock.prices",
           from = "2000-01-01",
           to   = "2019-12-31") %>%
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly", 
                 col_rename = "Rb")

baseline.returns
```

Now we will merge the datasets together to compute CAPM
```{r}
RaRb.maxret <- left_join(portfolio.returns.maxret, 
                                   baseline.returns,
                                   by = "date")

RaRb.minvar <- left_join(portfolio.returns.minvar, 
                                   baseline.returns,
                                   by = "date")

RaRb.qu <- left_join(portfolio.returns.qu, 
                                   baseline.returns,
                                   by = "date")

RaRb.qu2 <- left_join(portfolio.returns.qu2, 
                                   baseline.returns,
                                   by = "date")

RaRb.maxret
RaRb.minvar
RaRb.qu 
RaRb.qu2 
```

Now we will use the CAPM to evaluate the portfolios.

CAPM MaxRet Portfolio
```{r}
RaRb.maxret %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```

CAPM MinVar Portfolio
```{r}
RaRb.minvar %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```

CAPM Quadratic Utility Portfolio risk_aversion=1
```{r}
RaRb.qu %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```

CAPM Quadratic Utility Portfolio risk_aversion=5
```{r}
RaRb.qu2 %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```




### Visualizing Returns of Portfolios and baseline

To end this exercise, we will plot the returns of the portfolios and the returns of the baseline S&P 500.

First we need the baseline returns in xts format.
```{r}
baseline.returns.xts <- baseline.returns%>%
                      tk_xts(date_var = date, silent = TRUE)

baseline.returns.xts
```


Now we can plot profit and loss over time 
```{r}

bs.portf.maxret <- cbind(portfolio.returns.maxret$Ra, baseline.returns.xts)

charts.PerformanceSummary(bs.portf.maxret, main = "P/L Over Time MaxRet", )

```

```{r}

bs.portf.minvar <- cbind(portfolio.returns.minvar$Ra, baseline.returns.xts)

charts.PerformanceSummary(bs.portf.minvar, main = "P/L Over Time Minvar", )

```

```{r}

bs.portf.qu <- cbind(portfolio.returns.qu$Ra, baseline.returns.xts)

charts.PerformanceSummary(bs.portf.qu, main = "P/L Over Time RA=1", )

```


```{r}

bs.portf.qu2 <- cbind(portfolio.returns.qu2$Ra, baseline.returns.xts)

charts.PerformanceSummary(bs.portf.qu2, main = "P/L Over Time RA=5", )

```

The selection of the portfolio depends on the investor. All our portfolios perfom better than the S&P 500 index but three of the portfolios are riskier than the index. The Minimum Variance Portfolio has a positiv alpha and a beta under 1 so it is better in two categories, so we choose the Minimum Variance Portfolio, even though the other 3 portfolios have a higher alpha.


# Exercise 2: Custom moments function

Check `vignette("custom_moments_objectives")` to implement a variety of robust covariance matrix estimates (see `?MASS::cov.rob`, `?PerformanceAnalytics::ShrinkageMoments` and maybe `?PerformanceAnalytics::EWMAMoments` - the latter one only for backtesting) for the minimum variance and quadratic utility portfolios. Plot the different Efficient frontiers, optimal portfolios and weights and visualize the different covariances. Also make yourselves comfortable with cleaning outliers from your timeseries via `return.Clean()`.


Construct initial portfolio with basic constraints.
```{r}
portf.minvar <- portfolio.spec(assets=stocks.selection$symbol)%>%
  add.constraint(type="full_investment") %>%
  add.constraint(type="long_only") %>%
  add.objective(type="risk", name="StdDev")

portf.minvar
```

Next, we create initial maximum quadratic utility portfolio.
```{r}
portf.maxq <- portfolio.spec(assets=stocks.selection$symbol)%>%
  add.constraint(type="full_investment") %>%
  add.constraint(type="long_only") %>%
  add.objective(type="quadratic_utility", risk_aversion = 0.25)

portf.maxq
```


function to estimate covariance matrix with cov.rob for minimum variance

```{r}

sigma.robust <- function(R){
    require(MASS)
    out <- list()
    out$sigmarob <- cov.rob(R, method="mcd")$cov
    return(out)
}

sigmarob <- sigma.robust(stocks.returns.xts)$sigmarob

sigmarob
```

**function to estimate covariance matrix with ShrinkageMoments for minimum variance** 
Definition of Shrinkeage
Shrinkage is where extreme values in a sample are “shrunk” towards a central value, like the sample mean.

```{r}

sigma.robust.shrink <- function(R){
    targets <- c(1,3,4)
    f <- rowSums(stocks.returns.xts)
    out <- list()
    out$sigmashrink <- M2.shrink(stocks.returns.xts, targets, f)$M2sh
    return(out)
}
sigma.shrink <- sigma.robust.shrink(stocks.returns.xts)$sigmashrink
sigma.shrink
```

**Optimize portfolios**
Now we can use the custom moment function in optimize.portfolio to estimate the solution
to the minimum standard deviation portfolio.

Here we extract the weights and compute the portfolio standard deviation to verify that the
the robust estimate of the covariance matrix was used in the optimization.

```{r message=FALSE, warning=FALSE}
#portfolio moment "cov.rob"

##minimum variance portfolio
opt.sd.minv <- optimize.portfolio(stocks.returns.xts, portf.minvar, 
                                  optimize_method="ROI", 
                                  momentFUN="sigma.robust", 
                                  trace = TRUE)

##maximum quadratic utility portfolio
opt.sd.maxq <- optimize.portfolio(stocks.returns.xts, portf_maxqua, 
                                  optimize_method="ROI", 
                                  momentFUN="sigma.robust", 
                                  trace = TRUE)

#portfolio moment "ShrinkeageMoments"

##minimum variance portfolio
opt.sd.minv.shrink <- optimize.portfolio(stocks.returns.xts, portf_minvar,
                                         optimize_method="ROI",
                                         momentFUN="sigma.robust.shrink", 
                                         trace = TRUE)

##maximum quadratic utility portfolio
opt.sd.maxq.shrink <- optimize.portfolio(R=stocks.returns.xts,
                                         portfolio=portf.maxq, 
                                         optimize_method="ROI",
                                         momentFUN="sigma.robust.shrink", 
                                         trace = TRUE)

weights <- extractWeights(opt.sd.minv)
sigmarob <- sigma.robust(stocks.returns.xts)$sigmarob
sqrt(t(weights) %*% sigmarob %*% weights)

extractObjectiveMeasures(opt.sd.minv)$StdDev

opt.sd.minv
```

**Plot the covariance matrix from cov.rob**
```{r echo = FALSE}
chart.Correlation(sigmarob, histogram = TRUE)

```

Plot the covariance matrix from shrink
```{r echo = FALSE}
chart.Correlation(sigma.shrink, histogram = TRUE)
```
Plot the Portfolios
```{r}
plot(opt.sd.minv, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02))


plot(opt.sd.minv.shrink, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02))

plot(opt.sd.maxq, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.05))


plot(opt.sd.maxq.shrink, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.05))
```
**Chart Efficient Frontiert for the minimum variance Portfolio**
```{r echo = FALSE}
prt_eff_minv <- create.EfficientFrontier(R=stocks.returns.xts, 
                                         portfolio=portf.minvar, 
                                         type="mean-StdDev", 
                                         match.col = "StdDev")

chart.EfficientFrontier(prt_eff_minv, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)

chart.EF.Weights(prt_eff_minv, 
                 colorset=rainbow(n = length(stocks.selection$symbol)), 
                 match.col="StdDev", 
                 cex.lab = 1, 
                 main = "StdDev")
```

Chart Efficient Frontiert for the quadratic utility Portfolio
```{r echo = FALSE}
prt_eff_maxq <- create.EfficientFrontier(R=stocks.returns.xts, 
                                         portfolio=portf.maxq, 
                                         type="mean-StdDev", 
                                         match.col = "StdDev")

chart.EfficientFrontier(prt_eff_maxq, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)

chart.EF.Weights(prt_eff_maxq, colorset=rainbow(n = length(stocks.selection$symbol)),
                 match.col="StdDev", cex.lab = 1, main = "StdDev")
```


## Exercise 3: Regime Switching

Have a look at `demo(regime_switching)` and estimate and rebalance portfolios based on 2/3 regimes. Can you plot the regimes over time?

```{r load_packs}
pacman::p_load(tidyverse,tidyquant,PortfolioAnalytics,nloptr)
```




```{r}
# Load package and data.
library(PortfolioAnalytics)
```


```{r}
# Get monthly stock returns from the S&P500
monthly_returnsSP500 <- "^GSPC" %>%
  tq_get(get = "stock.prices", from = "2000-01-01", to = "2019-08-31") %>%
  tq_transmute(adjusted, periodReturn, period = "monthly", col_rename = "returns SP500")
monthly_returnsSP500
```

```{r}
# Calculate the rolling mean monthly

rollmeanSP500 <- rollmean(monthly_returnsSP500[, "returns SP500"], 2)
rollmeanSP500

```

```{r}
vector <- c(rollmeanSP500)

#2=good economy, 1=bad economy
regime1or2 <-as.numeric(vector>0)+1
regime1or2
```


```{r}
SP500dates <- select(monthly_returnsSP500,date)


#regime 1 is bad market phase (1) and regime 2 is good market phase (2)
data_frame <- data.frame(SP500dates[2:236,], regime1or2)
data_frame





```

```{r}
regime_xts <-data_frame %>%
  select(date, regime1or2)%>%
  timetk::tk_xts(silent = TRUE)
regime_xts
```


```{r}
stockselection <- c("ABCB", "AAPL", "ACLS", "ADBE", "ADTN", "AEHR", "AEIS", "AHPI", "AKAM", "AMZN")

# Get the prices of the stocks to transmute it to returns

stock.prices <- stockselection %>%
  tq_get(get  = "stock.prices", from = "2000-01-01",to = "2018-08-31") %>%
  group_by(symbol)

# Create monthly returns
stock.returns.monthly <- stock.prices %>%  
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns"
               )

## Make a tibble with dates and returns for all stocks
## Make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts (necessary for Portfolioanalytics)

R <- pivot_wider(stock.returns.monthly,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>% 
  timetk::tk_xts(date_var = date, silent = TRUE)



colnames(R) <- c("ABCB", "AAPL", "ACLS", "ADBE", "ADTN", "AEHR", "AEIS", "AHPI", "AKAM", "AMZN")
funds <- colnames(R)



R %>% head()
```


```{r}
# Construct portfolio for regime 1 - bad economy.
## Here, the first regime is considered with a risk approach (Mean-ES portfolio and other constraints)   --> we optimize ES

### Es = Conditional Value at risk: considers losses that exceed the value-at-risk and determines their average amount. 

port1 <- portfolio.spec(funds)
port1 <- add.constraint(port1, "weight_sum", min_sum=0.99, max_sum=1.01)
port1 <- add.constraint(port1, "box", min=0.05, max=0.5)
port1 <- add.objective(port1, type="risk", name="ES", arguments=list(p=0.9))
port1 <- add.objective(port1, type="risk_budget", name="ES", 
                       arguments=list(p=0.9), max_prisk=0.5)

```


```{r}
## Construct portfolio for regime 2 - good economy.
### Here regime 2 is a regime based on standard investment with volatility - here we used the standard deviation --> we optimize Stdev

port2 <- portfolio.spec(funds)
port2 <- add.constraint(port2, "weight_sum", min_sum=0.99, max_sum=1.01)
port2 <- add.constraint(port2, "box", min=0, max=0.6)
port2 <- add.objective(port2, type="risk", name="StdDev")
port2 <- add.objective(port2, type="risk_budget", name="StdDev", max_prisk=0.5)
```


```{r}
# Combine the portfolios.
portfolios <- combine.portfolios(list(port1, port2))


# Now we construct the regime model and corresponding portfolios to use for each regime.

## we merge the portfolios and the regimes (because we cannot merge every single portfolio with the regimes)

regime.port <- regime.portfolios(regime_xts, portfolios)
regime.port
```
```{r}
### This optimization should result in out portfolio for regime 2
opt1 <- optimize.portfolio(R[1:(nrow(R)-1)], regime.port, 
                           optimize_method="DEoptim", 
                           search_size=2000, 
                           trace=TRUE)
```
```{r}
opt1
opt1$regime
```


```{r}
### This optimization should result in out portfolio for regime 1
opt2 <- optimize.portfolio(R[1:(nrow(R)-1)], regime.port, 
                           optimize_method="DEoptim", 
                           search_size=2000, 
                           trace=TRUE)
```
```{r}
opt2
opt2$regime
```


```{r}
# We can extract which regime portfolio we optimized with at each rebalance date.  

## If there are structural changes in the data series, maybe a date fits better in the other regime now
lapply(opt.rebal$opt_rebalancing, function(x) x$regime)
```


```{r}
## Extract the optimal weights at each rebalance date.
wt <- extractWeights(opt.rebal)
wt
```


```{r}
## Extract the objective measures.
obj <- extractObjectiveMeasures(opt.rebal)
str(obj)
obj
```



```{r}
## Extract the stats.
xt <- extractStats(opt.rebal)
str(xt)
```


```{r}
### Note that this returns a list of N elements for N regimes. We may have different objectives and/or a different number of objectives which makes returning a single xts object difficult/

# Extract the optimal weights at each rebalance date.
chart.Weights(opt.rebal, colorset=rainbow10equal)

wt
```

```{r}
# Chart the risk contribution for regime 1

chart.RiskBudget(opt.rebal, match.col="ES", risk.type="percentage", 
                 regime=1, colorset=rainbow10equal)

opt2
```


```{r}
# Chart the risk contribution for regime 2
chart.RiskBudget(opt.rebal, match.col="StdDev", risk.type="percentage", 
                 regime=2, colorset=rainbow10equal)

opt1
```

```{r}
# Chart the risk contribution for regime 2
chart.RiskBudget(opt.rebal, match.col="StdDev", risk.type="percentage", 
                 regime=2, colorset=rainbow10equal)

opt1
```







## Exercise 4: Single Index-Model

Now we are going to estimate the Portfolio Input Parameters with the Single-Index Model. Use your ten assets and additionally choose the S&P500 as index (same returns etc).

a) Regress all stocks on the index. Show alpha, beta and residual variance. Calculate systematic and firm-specific risk. Are there any significant alphas? (You should double check with the appropriate `PerformanceAnalytics` Functions)





b) Extract the betas and calculate systematic and unsystematic risk, derive the whole covariance matrix. To do this you can use _CH15_Factor_Modfels_for_Asset_Returns.pdf (15.3.1)_ and the code
implemented in the function sharpeFactorEstimator that you find [here](http://financewithr.blogspot.com/2013/06/portfolio-optimization-using-single.html) (please do not just copy everything, but try to understand what you are doing, e.g. check why and if G.hat has the same values as found by the multivariate regression).









c) Now use the _custom-moments_ functions from Exercise 2 to implement the single-factor model into the portfolio optimization framework and plot the efficient frontier using the parameters estimated by the single factor model next to the EF of the full-covariance model. Calculate MVP, TP etc. and work out the differences in weights, portfolio return and portfolio risk.







### Get the packages 

```{r echo=FALSE}
install.packages("fEcofin", repos="http://R-Forge.R-project.org")

library(fEcofin)
library(fPortfolio)

```


Now we are going to estimate the Portfolio Input Parameters with the Single-Index Model. Use your ten assets and additionally choose the S&P500 as index (same returns etc).

```{r}

stockselection_4 <- c("MO", "MDLZ", "PEP", "PM", "KHC", "T","WMT", "MSFT","BAC","PG", "^GSPC")
stockselection_4

# Presettings
n <- length(stockselection_4)

#Get the prices of the stocks
stock.prices_4 <- stockselection_4 %>%
  tq_get(get  = "stock.prices", from = "2015-07-06",to = Sys.Date( )) %>% #First trade day of KHC
  group_by(symbol)

# Output the first two entries of each stock!
stock.prices_4 %>% slice(1:2) 
stock.prices_4 %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  ggtitle("Price chart for all stocks - all in one")

# Plotting the stock prices in each frame
stock.prices_4 %>%
  ggplot(aes(x = date, y = adjusted)) +
  geom_line() +
  facet_wrap(~symbol, scales = "free_y") +
  theme_classic() +
  labs(x = "Date", y = "Price") +
  ggtitle("Price chart all stocks - in each frame") 


```

```{r}

# Create monthly returns by the tq_transmute() = adds new variables to an existing tibble;
stock.returns.monthly_4 <- stock.prices_4 %>%  
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns")

# Output the first two entries of each stock!
stock.returns.monthly_4 %>% slice(1:2) 

# Make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts
stock.returns.monthly_xts_4 <- pivot_wider(stock.returns.monthly_4,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>%
   tk_xts(date_var = date, silent = TRUE)

# Output the first entries (simple returns from adjusted prices) of each stock!
stock.returns.monthly_xts_4[1]

# Plotting a performance summary chart 
charts.PerformanceSummary(stock.returns.monthly_xts_4, 
                          main="Performance summary")

```

a) Regress all stocks on the index. Show alpha, beta and residual variance. Calculate systematic and firm-specific risk. Are there any significant alphas? (You should double check with the appropriate `PerformanceAnalytics` Functions)

```{r}

#Regress all stocks on the index
alpha.Stocks <- CAPM.alpha(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], Rf = 0)
beta.Stocks <- CAPM.beta(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], Rf = 0)
StdDev.Index <- StdDev(R = stock.returns.monthly_xts_4[,n],
                   clean = "none",
                   method = "pearson")

lm(stock.returns.monthly_xts_4[,-n] ~ stock.returns.monthly_xts_4[,n])

for(i in 1:n) {
plot.default(x = stock.returns.monthly_xts_4[, n], 
            y = stock.returns.monthly_xts_4[, i], 
            main = stockselection_4[i], 
            xlab = "Index Returns", 
            ylab = "Stock Returns", 
            abline(lm(stock.returns.monthly_xts_4[, i] ~ stock.returns.monthly_xts_4[, n])))
}

#Calculate systematic (Market-Specific) Risk by mulitplying Variance (StdDev^2) of the S&P500 and the Beta^2 of each stock
sys.risk <- SystematicRisk(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], Rf = 0)
sys.risk

#Calculate Firm-specific Risk / Residual Variance
firm.specific.risk <- SpecificRisk(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], Rf = 0)
firm.specific.risk

#Summary
summary.SFM <- table.SFM(Ra = stock.returns.monthly_xts_4[,-n], Rb = stock.returns.monthly_xts_4[,n], scale = NA, Rf = 0, digits = 6)
summary.SFM

```

b) Extract the betas and calculate systematic and unsystematic risk, derive the whole covariance matrix. To do this you can use _CH15_Factor_Models_for_Asset_Returns.pdf (15.3.1)_ and the code
implemented in the function sharpeFactorEstimator that you find [here](http://financewithr.blogspot.com/2013/06/portfolio-optimization-using-single.html) (please do not just copy everything, but try to understand what you are doing, e.g. check why and if G.hat has the same values as found by the multivariate regression).

```{r}


#Calculate Beta of Portfolio by average each stocks beta
beta.portfolio <- mean(beta.Stocks)
beta.portfolio

#Calculate systematic (Market-Specific) Risk of portfolio
sys.risk.portfolio <- mean(sys.risk)
sys.risk.portfolio

#Calculate unsystematic risk by calculating the mean of the firm-specific risk
unsys.risk.portfolio <- mean(firm.specific.risk)
unsys.risk.portfolio

#Calculate Covariance Matrix
stock.returns.monthly_data <- as.data.frame((stock.returns.monthly_xts_4))
head(stock.returns.monthly_data)

returns <- as.timeSeries(stock.returns.monthly_data[,-n])
factors <- as.vector(as.timeSeries(stock.returns.monthly_data)[,n])

names(data)
data <- returns
attr(data, "factors") <- factors
nScenarios <- nrow(data)

X.mat <- cbind(rep(1, times=nScenarios), factors)

G.hat <- solve(qr(X.mat), data) #G.hat is alpha

beta.hat <- G.hat[2, ] #is beta

eps.hat <- data - X.mat %*% G.hat

diagD.hat <- diag(crossprod(eps.hat) / (nScenarios-2))

cov.si = var(factors)*(beta.hat%o%beta.hat) + diag(diagD.hat)

cov.si

```

c) Now use the _custom-moments_ functions from Exercise 2 to implement the single-factor model into the portfolo optimization framework and plot the efficient frontier using the parameters estimated by the single factor model next to the EF of the full-covariance model.

```{r}

#Function to implement single-factor model into the portfolio optimization framework

returns <- as.timeSeries(stock.returns.monthly_xts_4)
names(data)
data <- returns[, -c(n)]
factors <- returns[, n]
attr(data, "factors") <- factors

# Sharpe's Single Index Factor Model:
sharpeFactorEstimator <- 
function(x, spec=NULL, ...)
{
    # Sharpe Single Index Model:
    data <- getDataPart(x)
    factors <- attr(x, "factors")
    nScenarios <- nrow(data)
    X.mat <- cbind(rep(1, times=nScenarios), factors)
    G.hat <- solve(qr(X.mat), data)
    beta.hat <- G.hat[2, ]
    eps.hat <- data - X.mat %*% G.hat
    diagD.hat <- diag(crossprod(eps.hat) / (nScenarios-2))
    mu <- G.hat[1, ] + G.hat[2, ] * colMeans(factors)  
    Sigma <- var(factors)[[1]] * (beta.hat %o% beta.hat) + diag(diagD.hat)
    
    # Return Value:
    list(mu = mu, Sigma = Sigma)
}

spec <- portfolioSpec()
setEstimator(spec) <- "sharpeFactorEstimator"
sharpe <- portfolioFrontier(data, spec)

#Chart the efficient frontier using the parameters estimated by the single factor model
sharpe_1 <- portfolioFrontier(data)
tailoredFrontierPlot(sharpe_1)
points(frontierPoints(sharpe), col = "steelblue")


#Chart Efficient Frontier minimum variance
prt_eff_minv <- create.EfficientFrontier(R=stock.returns.monthly_xts_4, portfolio=SD.port.minv, type="mean-StdDev", match.col = "StdDev")
chart.EfficientFrontier(prt_eff_minv, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)

```

Calculate MVP, TP etc. and work out the differences in weights, portfolio return and portfolio risk.

```{r}

#weights, portfolio return and portfolio risk of MVP
opt.sd.minv.shrink
plot(opt.sd.minv.shrink, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02)) 


#weights, portfolio return and portfolio risk of QUP
opt.sd.maxq.shrink
plot(opt.sd.maxq.shrink, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(-0.08,0.05))


#weights, portfolio return and portfolio risk of tangency portfolio (Highest risk/return ratio)
weight_tp_sharpe <- sharpe_1@portfolio@portfolio[["weights"]][26, ]
return_tp_sharpe <- sharpe_1@portfolio@portfolio[["targetReturn"]][26, ]
risk_tp_sharpe <- sharpe_1@portfolio@portfolio[["targetRisk"]][26, ]

weight_tp_sharpe
return_tp_sharpe
risk_tp_sharpe


````









