---
title: "Portfoliomanagement and Financial Analysis - Assignment 4"
subtitle: "Submit until Monday 2019-10-07, 13:00"
author: "Caliskan, Mevhibe"
output: html_notebook
---

```{r load_packs}
pacman::p_load(tidyverse,tidyquant,PortfolioAnalytics,nloptr)
```

**Please** remember to put your assignment solutions in `rmd` format using **many** chunks and putting readable text in between, similar to my examples given in Research Methods and Assignment 1! Also, each student has to select his own set of 10 stocks having data available as of `2000-01-01`. Select by Sharpe-ratio, dominance or any other method (e.g. matching your first name).

For all exercises: Please use the Assignment-Forum to post your questions, I will try my best to help you along!

## Exercise 1: Rebalancing

Have a look at `vignette("ROI_vignette")` and the `optimize.portfolio.rebalancing` command. Use your dataset to compute 


```{r}
library(timetk)

SP500 <- tq_index("SP500")
NASDAQ <- tq_exchange("NASDAQ")
NYSE <- tq_exchange("NYSE") 

stocks.selection <- SP500 %>% 
  inner_join(rbind(NYSE,NASDAQ) %>% select(symbol,last.sale.price,market.cap,ipo.year),by=c("symbol")) %>%
  filter(ipo.year<2000&!is.na(market.cap)) %>% 
  arrange(desc(weight)) %>% 
  slice(1:10)

stocks.selection
```
These are the returns of the selected stocks.

```{r}
stocks.returns <- stocks.selection$symbol %>%
    tq_get(get  = "stock.prices",
           from = "2000-01-01",
           to   = "2019-12-31") %>%
    group_by(symbol) %>%
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly")

stocks.returns
```
Stock returns as time series

````{r}
stocks.returns.xts <- stocks.returns%>%
                      subset( select = c(symbol,date, monthly.returns)) %>% 
                      pivot_wider(names_from = symbol, 
                                  values_from = monthly.returns) %>% 
                      tk_xts(date_var = date, silent = TRUE)

stocks.returns.xts
```


a) Mean-Return (tangency portfolio)



Create Portfolio Object, add constraints, add objective

```{r}
portf_maxret <- portfolio.spec(assets=stocks.selection$symbol)%>%
  add.constraint(type="full_investment") %>%
  add.constraint(type="long_only") %>%
  add.objective(type="return", name="mean")
portf_maxret
```


Run the optimization 

```{r}

opt_maxret <- optimize.portfolio(R=stocks.returns.xts, portfolio=portf_maxret,
                                 optimize_method="ROI", trace=TRUE)
opt_maxret

```

```{r}

plot(opt_maxret, chart.assets=TRUE, main="Maximum Return", 
      xlim=c(0.05,0.35), ylim=c(0,0.04))


```


```{r}

maxret.ef <- create.EfficientFrontier(R=stocks.returns.xts, 
                                       portfolio=portf_maxret, 
                                       type="mean-StdDev")

chart.EfficientFrontier(maxret.ef, match.col="StdDev", type="l")



```



Backtesting:  Quarterly rebalancing with 5 year training period

```{r}

bt_maxret <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_maxret,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period=60)

bt_maxret 


```
```{r}
chart.Weights(bt_maxret, main = "Rebalanced Weights Over Time",
              colorset=rainbow(n = length(stocks.selection$symbol)))
```


Backtesting:  Quarterly rebalancing with 5 year training period and 4 year rolling window

```{r}

bt_maxret2 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_maxret,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period= 60,
                                            rolling_window = 48)
bt_maxret2

```
```{r}
chart.Weights(bt_maxret2, main = "Rebalanced Weights Over Time",
              colorset=rainbow(n = length(stocks.selection$symbol)))
```              


b) Minimum-Variance

Create portfolio object, add constraints, add objective

```{r}

portf_minvar <- portfolio.spec(assets= stocks.selection$symbol) %>%
  add.constraint(type="full_investment") %>%
  add.constraint(type = "long_only") %>%
  add.objective(type="risk", name="var")
  
portf_minvar

```
Run the optimization

```{r}
opt_minvar <- optimize.portfolio(R = stocks.returns.xts,portfolio = portf_minvar,
                              optimize_method = "ROI", trace = TRUE)

opt_minvar
```
Plotting weights

```{r}

plot(opt_minvar, chart.assets=TRUE, main="Minimum Variance with Constraints", 
      xlim=c(0.05, 0.35), ylim=c(0, 0.05))


```

Backtesting: Quarterly rebalancing with 5 year training period

```{r}
bt_minvar <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_minvar,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period=60)
bt_minvar
```
```{r}
chart.Weights(bt_minvar, main = "Rebalanced Weights Over Time",
              colorset=rainbow(n = length(stocks.selection$symbol)))
```


Backtesting: Quarterly rebalancing with 5 year training period and 4 year rolling window

```{r}
bt_minvar2 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, 
                                            portfolio=portf_minvar,
                                            optimize_method="ROI",
                                            rebalance_on="quarters",
                                            training_period=60,
                                            rolling_window = 48)
bt_minvar2
```
```{r}
chart.Weights(bt_minvar2, main = "Rebalanced Weights Over Time",
              colorset=rainbow(n = length(stocks.selection$symbol)))
```



c) Maximum Quadratic Utility Portfolios

```{r}

portf_maxqua <- portfolio.spec(assets= stocks.selection$symbol) %>%
  add.constraint(type = "full_investment") %>%
  add.constraint(type = "long_only")
portf_maxqua
```

### With risk aversion = 1
Create return objective, risk aversion objective and combine

```{r}
ret_obj <- return_objective(name="mean")

var_obj <- portfolio_risk_objective(name="var", risk_aversion=1)

qu_obj <- list(ret_obj, var_obj)

qu_obj
```
Run the optimization

```{r}
opt_qu <- optimize.portfolio(R=stocks.returns.xts, portfolio=portf_maxqua,
                             objectives=qu_obj,
                             optimize_method="ROI",
                             trace=TRUE)
opt_qu
```
```{r}
plot(opt_qu, chart.assets=TRUE, main="Maximum Quadratic Utility", 
      xlim=c(0, 0.35), ylim=c(0,0.05))
```


Backtesting: Quarterly rebalancing with 5 year training period

```{r}
bt_qu <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60)

bt_qu

```
```{r}
chart.Weights(bt_qu, main = "Rebalanced Weights Over Time",
              colorset=rainbow(n = length(stocks.selection$symbol)))
```

### With risk aversion = 5

Create risk aversion objective and combine with return objective

```{r}
var_obj2 <- portfolio_risk_objective(name="var", risk_aversion=5)

qu_obj2 <- list(ret_obj, var_obj2)

qu_obj2
```
Run the optimization

```{r}
opt_qu2 <- optimize.portfolio(R=stocks.returns.xts, portfolio=portf_maxqua,
                             objectives=qu_obj2,
                             optimize_method="ROI",
                             trace=TRUE)
opt_qu2
```
```{r}
plot(opt_qu2, chart.assets=TRUE, main="Maximum Quadratic Utility", 
      xlim=c(0, 0.35), ylim=c(0,0.05))
```


Backtesting: Quarterly rebalancing with 5 year training period

```{r}
bt_qu3 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj2,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60)
bt_qu
```
```{r}
chart.Weights(bt_qu3, main = "Rebalanced Weights Over Time",
              colorset=rainbow(n = length(stocks.selection$symbol)))
```



Backtesting: Quarterly rebalancing with 5 year training period

```{r}
bt_qu4 <- optimize.portfolio.rebalancing(R=stocks.returns.xts, portfolio=portf_maxqua,
                                        objectives=qu_obj2,
                                        optimize_method="ROI",
                                        rebalance_on="quarters",
                                        training_period=60,
                                        rolling_window = 48)
bt_qu4
```
```{r}
chart.Weights(bt_qu4, main = "Rebalanced Weights Over Time",
              colorset=rainbow(n = length(stocks.selection$symbol)))
```


### Comparing Portfolio Performance with CAPM

weights of the portfolios

```{r}
wts_maxret <- extractWeights(opt_maxret)
wts_minvar <- extractWeights(opt_minvar)
wts_qu <- extractWeights(opt_qu)
wts_qu2 <- extractWeights(opt_qu2)


wts_maxret
wts_minvar
wts_qu 
wts_qu2
```
Portfolio Returns

```{r}
stocks.returns

portfolio.returns.maxret <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_maxret, 
                 col_rename  = "Ra")

portfolio.returns.maxret

portfolio.returns.minvar <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_minvar, 
                 col_rename  = "Ra")

portfolio.returns.minvar

portfolio.returns.qu <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_qu, 
                 col_rename  = "Ra")

portfolio.returns.qu

portfolio.returns.qu2 <- stocks.returns %>%
    tq_portfolio(assets_col  = symbol, 
                 returns_col = monthly.returns, 
                 weights     = wts_qu2, 
                 col_rename  = "Ra")

portfolio.returns.qu2
```
Baseline Returns S&P 500

```{r}
baseline.returns <- "^GSPC" %>%
    tq_get(get  = "stock.prices",
           from = "2000-01-01",
           to   = "2019-12-31") %>%
    tq_transmute(select     = adjusted, 
                 mutate_fun = periodReturn, 
                 period     = "monthly", 
                 col_rename = "Rb")

baseline.returns
```


Merging the datasets together to compute CAPM

```{r}
RaRb.maxret <- left_join(portfolio.returns.maxret, 
                                   baseline.returns,
                                   by = "date")

RaRb.minvar <- left_join(portfolio.returns.minvar, 
                                   baseline.returns,
                                   by = "date")

RaRb.qu <- left_join(portfolio.returns.qu, 
                                   baseline.returns,
                                   by = "date")

RaRb.qu2 <- left_join(portfolio.returns.qu2, 
                                   baseline.returns,
                                   by = "date")

RaRb.maxret
RaRb.minvar
RaRb.qu 
RaRb.qu2 
```
CAPM MaxRet Portfolio

```{r}
RaRb.maxret %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```


CAPM MinVar Portfolio

```{r}
RaRb.minvar %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```

CAPM Quadratic Utility Portfolio risk_aversion=1

```{r}
RaRb.qu %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```

CAPM Quadratic Utility Portfolio risk_aversion=5

```{r}
RaRb.qu2 %>%
    tq_performance(Ra = Ra, Rb = Rb, 
    performance_fun = table.CAPM)
```
### Visualizing Returns of Portfolios and baseline

Baseline Returns in xts format

```{r}
baseline.returns.xts <- baseline.returns%>%
                      tk_xts(date_var = date, silent = TRUE)

baseline.returns.xts
```
Profit and Loss over time 
```{r}

bs.portf.maxret <- cbind(portfolio.returns.maxret$Ra, baseline.returns.xts)

charts.PerformanceSummary(bs.portf.maxret, main = "P/L Over Time MaxRet", )

```
```{r}

bs.portf.minvar <- cbind(portfolio.returns.minvar$Ra, baseline.returns.xts)

charts.PerformanceSummary(bs.portf.minvar, main = "P/L Over Time Minvar", )

```
```{r}

bs.portf.qu <- cbind(portfolio.returns.qu$Ra, baseline.returns.xts)

charts.PerformanceSummary(bs.portf.qu, main = "P/L Over Time", )

```
```{r}

bs.portf.qu2 <- cbind(portfolio.returns.qu2$Ra, baseline.returns.xts)

charts.PerformanceSummary(bs.portf.qu2, main = "P/L Over Time", )

```


checking for a variety of constraints (constraints that can be computed with the `ROI`-solver) and different rebalancing periods (as well as rolling windows/training periods) to find, what might deliver you the best portfolios performance (use appropriate statistics to decide on that).



## Exercise 2: Custom moments function

Check `vignette("custom_moments_objectives")` to implement a variety of robust covariance matrix estimates (see `?MASS::cov.rob`, `?PerformanceAnalytics::ShrinkageMoments` and maybe `?PerformanceAnalytics::EWMAMoments` - the latter one only for backtesting) for the minimum variance and quadratic utility portfolios. Plot the different Efficient frontiers, optimal portfolios and weights and visualize the different covariances. Also make yourselves comfortable with cleaning outliers from your timeseries via `return.Clean()`.

```{r}
require(timetk)

stock.prices <- stocks.selection %>%
  tq_get(get  = "stock.prices", from = "2000-01-01",to = "2018-08-31") %>%
  group_by(symbol)

stock.returns.monthly <- stock.prices %>%  
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns"
               )

stock.returns.monthly

```

Now, we make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts. The method "Return_Clean()" cleans the returns from outliers.

```{r}
stock.returns.monthly_xts_with_outliers <- pivot_wider(stock.returns.monthly,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>% 
  tk_xts(date_var = date, silent = TRUE)

stock.returns.monthly_xts <- Return.clean(stock.returns.monthly_xts_with_outliers, method = "boudt", alpha = 0.01)
options(max.print=60)
stock.returns.monthly_xts

```

Now, we create the initial minimum variance portfolio

```{r}
require(PortfolioAnalytics)
require(DEoptim)
require(ROI)
require(ROI.plugin.glpk)
require(ROI.plugin.quadprog)

# Construct initial portfolio with basic constraints.

init.port.minv <- portfolio.spec(assets=colnames(stock.returns.monthly_xts),category_labels = stocks.selection)

init.port.minv <- add.constraint(portfolio=init.port.minv, type="full_investment")

init.port.minv <- add.constraint(portfolio=init.port.minv, type="long_only")

#Portfolio with standard deviation as an objective

SD.port.minv <- add.objective(portfolio=init.port.minv, type="risk", name="StdDev")

#init.port.minv

SD.port.minv
```

Next, we create initial maximum quadratic utility portfolio.

```{r}
# Construct initial portfolio with basic constraints.
init.port.maxq <- portfolio.spec(assets=colnames(stock.returns.monthly_xts),category_labels = stocks.selection)

#init.port.maxq <- add.constraint(init.port.maxq, type = "box", min = 0, max = 1)
init.port.maxq <- add.constraint(portfolio=init.port.maxq, type="full_investment")
init.port.maxq <- add.constraint(portfolio=init.port.maxq, type="long_only")

#Portfolio with standard deviation as an objective
#SD.port.maxq <- add.objective(portfolio=init.port.maxq, type="return", name="mean")
#SD.port.maxq <- add.objective(portfolio=SD.port.maxq, type="risk", name="var", risk_aversion=0.25)
SD.port.maxq <- add.objective(portfolio=init.port.maxq, type="quadratic_utility", risk_aversion=0.25)

SD.port.maxq
```

function to estimate covariance matrix with cov.rob for minimum variance

```{r}

sigma.robust <- function(R){
    require(MASS)
    out <- list()
    out$sigmarob <- cov.rob(R, method="mcd")$cov
    return(out)
}

sigmarob <- sigma.robust(stock.returns.monthly_xts)$sigmarob

sigmarob
```

function to estimate covariance matrix with ShrinkageMoments for minimum variance

```{r}

sigma.robust.shrink <- function(R){
    targets <- c(1,3,4)
    f <- rowSums(stock.returns.monthly_xts)
    out <- list()
    out$sigmashrink <- M2.shrink(stock.returns.monthly_xts, targets, f)$M2sh
    return(out)
}
sigma.shrink <- sigma.robust.shrink(stock.returns.monthly_xts)$sigmashrink
sigma.shrink
```
Optimize portfolios

```{r message=FALSE, warning=FALSE}
#portfolio moment "cov.rob"

##minimum variance portfolio
opt.sd.minv <- optimize.portfolio(stock.returns.monthly_xts, SD.port.minv, optimize_method="ROI", momentFUN="sigma.robust", trace = TRUE)

##maximum quadratic utility portfolio
opt.sd.maxq <- optimize.portfolio(stock.returns.monthly_xts, SD.port.maxq, optimize_method="ROI", momentFUN="sigma.robust", trace = TRUE)

#portfolio moment "ShrinkeageMoments"

##minimum variance portfolio
opt.sd.minv.shrink <- optimize.portfolio(stock.returns.monthly_xts, SD.port.minv, optimize_method="ROI", momentFUN="sigma.robust.shrink", trace = TRUE)

##maximum quadratic utility portfolio
opt.sd.maxq.shrink <- optimize.portfolio(R=stock.returns.monthly_xts, portfolio=SD.port.maxq, optimize_method="ROI", momentFUN="sigma.robust.shrink", trace = TRUE)

weights <- extractWeights(opt.sd.minv)
sigmarob <- sigma.robust(stock.returns.monthly_xts)$sigmarob
sqrt(t(weights) %*% sigmarob %*% weights)

extractObjectiveMeasures(opt.sd.minv)$StdDev

opt.sd.minv
```

Plot the covariance matrix from cov.rob
```{r echo = FALSE}
chart.Correlation(sigmarob, histogram = TRUE)

```

Plot the covariance matrix from shrink

```{r echo = FALSE}
chart.Correlation(sigma.shrink, histogram = TRUE)
```

Plot the Portfolios

```{r}
plot(opt.sd.minv, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02))


plot(opt.sd.minv.shrink, risk.col="StdDev", return.col="mean", main="Minimum Variance Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.02))

plot(opt.sd.maxq, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.05))


plot(opt.sd.maxq.shrink, risk.col="StdDev", return.col="mean", main="Quadratic Utility Optimization shrink", chart.assets=TRUE, xlim=c(0, 0.2), ylim=c(0,0.05))
```

Chart Efficient Frontiert for the minimum variance Portfolio

```{r echo = FALSE}
prt_eff_minv <- create.EfficientFrontier(R=stock.returns.monthly_xts, portfolio=SD.port.minv, type="mean-StdDev", match.col = "StdDev")

chart.EfficientFrontier(prt_eff_minv, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)
chart.EF.Weights(prt_eff_minv, colorset=rainbow(n = length(stockselection)), match.col="StdDev", cex.lab = 1, main = "StdDev")
```

Chart Efficient Frontiert for the quadratic utility Portfolio

```{r echo = FALSE}
prt_eff_maxq <- create.EfficientFrontier(R=stock.returns.monthly_xts, portfolio=SD.port.maxq, type="mean-StdDev", match.col = "StdDev")
chart.EfficientFrontier(prt_eff_maxq, match.col="StdDev", type="b", rf=NULL, pch.assets = 1)
chart.EF.Weights(prt_eff_maxq, colorset=rainbow(n = length(stockselection)), match.col="StdDev", cex.lab = 1, main = "StdDev")
```

## Exercise 3: Regime Switching

Have a look at `demo(regime_switching)` and estimate and rebalance portfolios based on 2/3 regimes. Can you plot the regimes over time?


```{r}
# Load package and data.
library(PortfolioAnalytics)
```

```{r}
# Calculate the rolling mean monthly

rollmeanSP500 <- rollmean(stock.returns.monthly[, "Stock.returns"], 2)
rollmeanSP500

```

```{r}
vector <- c(rollmeanSP500)

#2=good economy, 1=bad economy
regime1or2 <-as.numeric(vector>0)+1
regime1or2
```
```{r}
SP500dates <- select(stock.returns.monthly,date)


#regime 1 is bad market phase (1) and regime 2 is good market phase (2)
data_frame <- data.frame(SP500dates[2:236,], regime1or2)
data_frame


```

```{r}
regime_xts <-data_frame %>%
  select(date, regime1or2)%>%
  timetk::tk_xts(silent = TRUE)
regime_xts
```

```{r}
## Make a tibble with dates and returns for all stocks
## Make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts (necessary for Portfolioanalytics)

R <- pivot_wider(stock.returns.monthly,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>% 
  timetk::tk_xts(date_var = date, silent = TRUE)



colnames(R) <- c("AAPL", "MSFT", "AMZN", "NVDA", "ADBE", "CSCO", "AMGN", "QCOM", "UPS", "ORCL")
funds <- colnames(R)



R %>% head()
```

```{r}
# Construct portfolio for regime 1 - bad economy.
## Here, the first regime is considered with a risk approach (Mean-ES portfolio and other constraints)   --> we optimize ES

### Es = Conditional Value at risk: considers losses that exceed the value-at-risk and determines their average amount. 

port1 <- portfolio.spec(funds)
port1 <- add.constraint(port1, "weight_sum", min_sum=0.99, max_sum=1.01)
port1 <- add.constraint(port1, "box", min=0.05, max=0.5)
port1 <- add.objective(port1, type="risk", name="ES", arguments=list(p=0.9))
port1 <- add.objective(port1, type="risk_budget", name="ES", 
                       arguments=list(p=0.9), max_prisk=0.5)

```


```{r}
## Construct portfolio for regime 2 - good economy.
### Here regime 2 is a regime based on standard investment with volatility - here we used the standard deviation --> we optimize Stdev

port2 <- portfolio.spec(funds)
port2 <- add.constraint(port2, "weight_sum", min_sum=0.99, max_sum=1.01)
port2 <- add.constraint(port2, "box", min=0, max=0.6)
port2 <- add.objective(port2, type="risk", name="StdDev")
port2 <- add.objective(port2, type="risk_budget", name="StdDev", max_prisk=0.5)
```

```{r}
# Combine the portfolios.
portfolios <- combine.portfolios(list(port1, port2))


# Now we construct the regime model and corresponding portfolios to use for each regime.

## we merge the portfolios and the regimes (because we cannot merge every single portfolio with the regimes)

regime.port <- regime.portfolios(regime_xts, portfolios)
regime.port
```

```{r}
### This optimization should result in out portfolio for regime 2
opt1 <- optimize.portfolio(R[1:(nrow(R)-1)], regime.port, 
                           optimize_method="DEoptim", 
                           search_size=2000, 
                           trace=TRUE)
```
```{r}
opt1
opt1$regime
```
```{r}
### This optimization should result in out portfolio for regime 1
opt2 <- optimize.portfolio(R[1:(nrow(R)-1)], regime.port, 
                           optimize_method="DEoptim", 
                           search_size=2000, 
                           trace=TRUE)
```
```{r}
opt2
opt2$regime
```


```{r}
# We can extract which regime portfolio we optimized with at each rebalance date.  

## If there are structural changes in the data series, maybe a date fits better in the other regime now
lapply(opt.rebal$opt_rebalancing, function(x) x$regime)
```


```{r}
## Extract the optimal weights at each rebalance date.
wt <- extractWeights(opt.rebal)
wt
```

```{r}
## Extract the objective measures.
obj <- extractObjectiveMeasures(opt.rebal)
str(obj)
obj
```

```{r}
## Extract the stats.
xt <- extractStats(opt.rebal)
str(xt)
```


```{r}
### Note that this returns a list of N elements for N regimes. We may have different objectives and/or a different number of objectives which makes returning a single xts object difficult/

# Extract the optimal weights at each rebalance date.
chart.Weights(opt.rebal, colorset=rainbow10equal)

wt
```
```{r}
# Chart the risk contribution for regime 1

chart.RiskBudget(opt.rebal, match.col="ES", risk.type="percentage", 
                 regime=1, colorset=rainbow10equal)

opt2
```


```{r}
# Chart the risk contribution for regime 2
chart.RiskBudget(opt.rebal, match.col="StdDev", risk.type="percentage", 
                 regime=2, colorset=rainbow10equal)

opt1
```

```{r}
# Chart the risk contribution for regime 2
chart.RiskBudget(opt.rebal, match.col="StdDev", risk.type="percentage", 
                 regime=2, colorset=rainbow10equal)

opt1
```



## Exercise 4: Single Index-Model

Now we are going to estimate the Portfolio Input Parameters with the Single-Index Model. Use your ten assets and additionally choose the S&P500 as index (same returns etc).

a) Regress all stocks on the index. Show alpha, beta and residual variance. Calculate systematic and firm-specific risk. Are there any significant alphas? (You should double check with the appropriate `PerformanceAnalytics` Functions)
b) Extract the betas and calculate systematic and unsystematic risk, derive the whole covariance matrix. To do this you can use _CH15_Factor_Modfels_for_Asset_Returns.pdf (15.3.1)_ and the code
implemented in the function sharpeFactorEstimator that you find [here](http://financewithr.blogspot.com/2013/06/portfolio-optimization-using-single.html) (please do not just copy everything, but try to understand what you are doing, e.g. check why and if G.hat has the same values as found by the multivariate regression).
c) Now use the _custom-moments_ functions from Exercise 2 to implement the single-factor model into the portfolio optimization framework and plot the efficient frontier using the parameters estimated by the single factor model next to the EF of the full-covariance model. Calculate MVP, TP etc. and work out the differences in weights, portfolio return and portfolio risk.

```{r}
# get monthly stock returns from the S&P500
monthly_returnsSP500 <- "^GSPC" %>%
  tq_get(get = "stock.prices", from = "2000-01-01", to = "2019-08-31") %>%
  tq_transmute(adjusted, periodReturn, period = "monthly", col_rename = "returns SP500")
monthly_returnsSP500
```
```{r}
#calculate the rolling mean monthly
rollmeanSP500 <- rollmean(monthly_returnsSP500[, "returns SP500"], 2)
rollmeanSP500
```

```{r}
vector <- c(rollmeanSP500)
#2=good economy, 1=bad economy
regime1or2 <-as.numeric(vector>0)+1
regime1or2
```
```{r}
SP500dates <- as.data.frame(monthly_returnsSP500$date)
#regime 1 is bad market phase (1) and regime 2 is good market phase (2)
data_frame <- data.frame("date" =SP500dates[2:236,], regime1or2)
data_frame
#convert it in xts format

regime_xts <- data_frame %>%
  data.frame[1] %>%
   tk_xts(date_var = data_frame$date, silent = TRUE)
regime_xts
```

```{r}
stockselection <- c("AAPL", "MSFT", "AMZN", "JPM", "AMGN", "CSCO", "GILD", "MCD", "QCOM", "WMT")
#Get the prices of the stocks to transmute it to returns
stock.prices <- stockselection %>%
  tq_get(get  = "stock.prices", from = "2000-01-01",to = "2018-08-31") %>%
  group_by(symbol)
#Create monthly returns
stock.returns.monthly <- stock.prices %>%  
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="monthly",
               type="arithmetic",
               col_rename = "Stock.returns"
               )
#make a tibble with dates and returns for all stocks
#Make 10 columns (each for every stock) with the simple returns from adjusted prices and convert to xts (necessary for Portfolioanalytics)
R <- pivot_wider(stock.returns.monthly,
                                                names_from = symbol,
                                                values_from = c(Stock.returns))%>% 
  tk_xts(date_var = date, silent = TRUE)
colnames(R) <- c("AAPL", "MSFT", "AMZN", "JPM", "AMGN", "CSCO", "GILD", "MCD", "QCOM", "WMT")
funds <- colnames(R)
#Amgen -bio technology
#Gild - pharmacy
#QCOM - tele communication 
R %>% head()
```

set.seed(123)
regime <- xts(sample(1:2, nrow(R), replace=TRUE, prob=c(0.3, 0.7)), index(R))
#Regimes are simply market periods with different parameters, such as good market phases (upswing) and bad (downturn). Regimes are different economic environments (e.g. good times and bad times). 
# We want  a different portfolio for good times and for bad times (and maybe - as a third regime - for times that are uncertain but neither really good nor bad). --> 
#here, they assign different dates to regime 1 or 2
#there are more in regime 2 because prob. of regime 1 is 30% and of regime 2 is 70%
regime
```
```{r}
#' Construct portfolio for regime 1 - bad economy.
#here, the first regime is considered with a risk approach (Mean-ES portfolio and other constraints)   --> we optimize ES
#Es = Conditional Value at risk: considers losses that exceed the value-at-risk and determines their average amount. 
port1 <- portfolio.spec(funds)
port1 <- add.constraint(port1, "weight_sum", min_sum=0.99, max_sum=1.01)
port1 <- add.constraint(port1, "box", min=0.05, max=0.5)
port1 <- add.objective(port1, type="risk", name="ES", arguments=list(p=0.9))
port1 <- add.objective(port1, type="risk_budget", name="ES", 
                       arguments=list(p=0.9), max_prisk=0.5)
```


```{r}
#' Construct portfolio for regime 2 - good economy.
#here regime 2 is a regime based on standard investment with volatility - here we used the standard deviation --> we optimize Stdev
port2 <- portfolio.spec(funds)
port2 <- add.constraint(port2, "weight_sum", min_sum=0.99, max_sum=1.01)
port2 <- add.constraint(port2, "box", min=0, max=0.6)
port2 <- add.objective(port2, type="risk", name="StdDev")
port2 <- add.objective(port2, type="risk_budget", name="StdDev", max_prisk=0.5)
```

```{r}
#' Combine the portfolios.
portfolios <- combine.portfolios(list(port1, port2))
#' Now we construct the regime model and corresponding portfolios to use for
#' each regime.
# we merge the portfolios and the regimes (becauase we cannot merge every single portfolio with the regimes)
regime.port <- regime.portfolios(regime_xts, portfolios)
regime.port
```

```{r}
#' This optimization should result in out portfolio for regime 2 (optimizing Stdev).
opt1 <- optimize.portfolio(R, regime.port, 
                           optimize_method="random", 
                           search_size=2000, 
                           trace=TRUE)
opt1
opt1$regime
```

```{r}
#' This optimization should result in out portfolio for regime 1.
opt2 <- optimize.portfolio(R[1:(nrow(R)-1)], regime.port, 
                           optimize_method="DEoptim", 
                           search_size=2000, 
                           trace=TRUE)
```

```{r}
opt2
opt2$regime
```

```{r}
#' Run optimization with rebalancing using our regime switching portfolio.
opt.rebal <- optimize.portfolio.rebalancing(R, regime.port,
                                            optimize_method="random", 
                                            rebalance_on="quarters", 
                                            training_period=130,
                                            search_size=2000, 
                                            trace=TRUE)
#' The print and summary methods work the same as they do for optimizations 
#' without regime switching.
opt.rebal
summary(opt.rebal)
```
```{r}
#' We can extract which regime portfolio we optimized with at each rebalance date.  
#if there are structural changes in the data series, maybe a date fits better in the other regime now
lapply(opt.rebal$opt_rebalancing, function(x) x$regime)
```


```{r}
#' Extract the optimal weights at each rebalance date.
wt <- extractWeights(opt.rebal)
wt
```

```{r}
#' Extract the objective measures*.
obj <- extractObjectiveMeasures(opt.rebal)
str(obj)
obj
```


```{r}
# Extract the stats*.
xt <- extractStats(opt.rebal)
str(xt)
```

```{r}
#' *
#' Note that this returns a list of N elements for N regimes. We may have 
#' different objectives and/or a different number of objectives which makes
#' returning a single xts object difficult/
#' Extract the optimal weights at each rebalance date.
chart.Weights(opt.rebal, colorset=rainbow10equal)
wt
```

```{r}
#' Chart the risk contribution for regime 1
# 
chart.RiskBudget(opt.rebal, match.col="ES", risk.type="percentage", 
                 regime=1, colorset=rainbow10equal)
opt2
```

```{r}
#' Chart the risk contribution for regime 2
chart.RiskBudget(opt.rebal, match.col="StdDev", risk.type="percentage", 
                 regime=2, colorset=rainbow10equal)
opt1
```

